# Financial Charts
## Time Series and Date Axess
### Time Series using Axes of type date
```r
library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', width = 900)


fig
```
```{r}
library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', width = 900)


fig
```
### Configuring Tick Labels
```r
library(tidyquant)
library(plotly)
tickers = c("GOOG", "AAPL", "AMZN", "META", "NFLX", "MSFT")
for (i in tickers){
  getSymbols(i,
             from = "2020-01-01",
             to = "2021-12-31")}
stock <- data.frame(GOOG$GOOG.Adjusted,
                    AAPL$AAPL.Adjusted,
                    AMZN$AMZN.Adjusted,
                    META$META.Adjusted,
                    NFLX$NFLX.Adjusted,
                    MSFT$MSFT.Adjusted)
stock$GOOG.Adjusted <- stock$GOOG.Adjusted/stock$GOOG.Adjusted[1]
stock$AAPL.Adjusted <- stock$AAPL.Adjusted/stock$AAPL.Adjusted[1]
stock$AMZN.Adjusted <- stock$AMZN.Adjusted/stock$AMZN.Adjusted[1]
stock$META.Adjusted <- stock$META.Adjusted/stock$META.Adjusted[1]
stock$NFLX.Adjusted <- stock$NFLX.Adjusted/stock$NFLX.Adjusted[1]
stock$MSFT.Adjusted <- stock$MSFT.Adjusted/stock$MSFT.Adjusted[1]
stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Dates, y = ~GOOG, name = 'GOOG')%>%
  add_trace(x = ~Dates, y = ~AAPL, name = 'AAPL')%>%
  add_trace(x = ~Dates, y = ~AMZN, name = 'AMZN')%>%
  add_trace(x = ~Dates, y = ~META, name = 'META')%>%
  add_trace(x = ~Dates, y = ~NFLX, name = 'NFLX')%>%
  add_trace(x = ~Dates, y = ~MSFT, name = 'MSFT')%>%
  layout(title = 'custom tick labels',legend=list(title=list(text='variable')),
         xaxis = list(dtick = "M1", tickformat="%b<br>%Y"), width = 1000)
options(warn = -1)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6')


fig
```
```{r}
library(tidyquant)
library(plotly)
tickers = c("GOOG", "AAPL", "AMZN", "META", "NFLX", "MSFT")
for (i in tickers){
  getSymbols(i,
             from = "2020-01-01",
             to = "2021-12-31")}
stock <- data.frame(GOOG$GOOG.Adjusted,
                    AAPL$AAPL.Adjusted,
                    AMZN$AMZN.Adjusted,
                    META$META.Adjusted,
                    NFLX$NFLX.Adjusted,
                    MSFT$MSFT.Adjusted)
stock$GOOG.Adjusted <- stock$GOOG.Adjusted/stock$GOOG.Adjusted[1]
stock$AAPL.Adjusted <- stock$AAPL.Adjusted/stock$AAPL.Adjusted[1]
stock$AMZN.Adjusted <- stock$AMZN.Adjusted/stock$AMZN.Adjusted[1]
stock$META.Adjusted <- stock$META.Adjusted/stock$META.Adjusted[1]
stock$NFLX.Adjusted <- stock$NFLX.Adjusted/stock$NFLX.Adjusted[1]
stock$MSFT.Adjusted <- stock$MSFT.Adjusted/stock$MSFT.Adjusted[1]
stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Dates, y = ~GOOG, name = 'GOOG')%>%
  add_trace(x = ~Dates, y = ~AAPL, name = 'AAPL')%>%
  add_trace(x = ~Dates, y = ~AMZN, name = 'AMZN')%>%
  add_trace(x = ~Dates, y = ~META, name = 'META')%>%
  add_trace(x = ~Dates, y = ~NFLX, name = 'NFLX')%>%
  add_trace(x = ~Dates, y = ~MSFT, name = 'MSFT')%>%
  layout(title = 'custom tick labels',legend=list(title=list(text='variable')),
         xaxis = list(dtick = "M1", tickformat="%b<br>%Y"), width = 1000)
options(warn = -1)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6')


fig
```
### Time Series with Range Selector Buttons
```r

library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F, title='Time Series with Range Slider and Selectors',
         xaxis = list(rangeslider = list(visible = T),
                      rangeselector=list(
                        buttons=list(
                          list(count=1, label="1m", step="month", stepmode="backward"),
                          list(count=6, label="6m", step="month", stepmode="backward"),
                          list(count=1, label="YTD", step="year", stepmode="todate"),
                          list(count=1, label="1y", step="year", stepmode="backward"),
                          list(step="all")
                        ))))
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', margin = 0.1, width = 900)
fig
```
```{r}

library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F, title='Time Series with Range Slider and Selectors',
         xaxis = list(rangeslider = list(visible = T),
                      rangeselector=list(
                        buttons=list(
                          list(count=1, label="1m", step="month", stepmode="backward"),
                          list(count=6, label="6m", step="month", stepmode="backward"),
                          list(count=1, label="YTD", step="year", stepmode="todate"),
                          list(count=1, label="1y", step="year", stepmode="backward"),
                          list(step="all")
                        ))))
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', margin = 0.1, width = 900)
fig
```

## Candlestick Charts
### Basic Candlestick
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
# basic example of ohlc charts
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic Candlestick Chart")

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
# basic example of ohlc charts
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic Candlestick Chart")

fig
```
### Customise the fig ure with Shapes and Annotations
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
)

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Apple Stock",
         annotations = a,
         shapes = l)

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
)

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Apple Stock",
         annotations = a,
         shapes = l)

fig
```

## OHLC Charts
### Basic OHLC Chart
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic OHLC Chart",
         xaxis = list(rangeslider = list(visible = F)))

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic OHLC Chart",
         xaxis = list(rangeslider = list(visible = F)))

fig
```
### Customise the Figure with Shapes and Annotations
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
          )

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
          )

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Custom Colors",
         annotations = a,
         shapes = l)

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
          )

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
          )

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Custom Colors",
         annotations = a,
         shapes = l)

fig
```


## Waterfall Charts
### Basic Waterfall Chart
```r
library(plotly)

x= list("Sales", "Consulting", "Net revenue", "Purchases", "Other expenses", "Profit before tax")
measure= c("relative", "relative", "total", "relative", "relative", "total")
text= c("+60", "+80", "", "-40", "-20", "Total")
y= c(60, 80, 0, -40, -20, 0)
data = data.frame(x=factor(x,levels=x),measure,text,y)

fig <- plot_ly(
  data, name = "20", type = "waterfall", measure = ~measure,
  x = ~x, textposition = "outside", y= ~y, text =~text,
  connector = list(line = list(color= "rgb(63, 63, 63)"))) 
fig <- fig %>%
  layout(title = "Profit and loss statement 2018",
        xaxis = list(title = ""),
        yaxis = list(title = ""),
        autosize = TRUE,
        showlegend = TRUE)

fig
```
```{r}
library(plotly)

x= list("Sales", "Consulting", "Net revenue", "Purchases", "Other expenses", "Profit before tax")
measure= c("relative", "relative", "total", "relative", "relative", "total")
text= c("+60", "+80", "", "-40", "-20", "Total")
y= c(60, 80, 0, -40, -20, 0)
data = data.frame(x=factor(x,levels=x),measure,text,y)

fig <- plot_ly(
  data, name = "20", type = "waterfall", measure = ~measure,
  x = ~x, textposition = "outside", y= ~y, text =~text,
  connector = list(line = list(color= "rgb(63, 63, 63)"))) 
fig <- fig %>%
  layout(title = "Profit and loss statement 2018",
        xaxis = list(title = ""),
        yaxis = list(title = ""),
        autosize = TRUE,
        showlegend = TRUE)

fig
```

## Funnel Charts
```r
# Need to install plotly from Github to get funnel plots
devtools::install_github("ropensci/plotly")
```
### Basic Funnel Plot
```r
library(plotly)

fig <- plot_ly() 
fig <- fig %>%
  add_trace(
  type = "funnel",
  y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
  x = c(39, 27.4, 20.6, 11, 2)) 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent")))

fig
```
```{r}
library(plotly)

fig <- plot_ly() 
fig <- fig %>%
  add_trace(
  type = "funnel",
  y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
  x = c(39, 27.4, 20.6, 11, 2)) 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent")))

fig
```
### Stacked Funnel Plot
```r
library(plotly)

fig <- plot_ly(
    type = "funnel",
    name = 'Montreal',
    y = c("Website visit", "Downloads", "Potential customers", "Requested price"),
    x = c(120, 60, 30, 20),
    textinfo = "value+percent initial") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Toronto',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
    x = c(100, 60, 40, 30, 20),
    textposition = "inside",
    textinfo = "value+percent previous") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Vancouver',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized"),
  x = c(90, 70, 50, 30, 10, 5),
  textposition = "outside",
  textinfo = "value+percent total") 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized")))

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = "funnel",
    name = 'Montreal',
    y = c("Website visit", "Downloads", "Potential customers", "Requested price"),
    x = c(120, 60, 30, 20),
    textinfo = "value+percent initial") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Toronto',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
    x = c(100, 60, 40, 30, 20),
    textposition = "inside",
    textinfo = "value+percent previous") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Vancouver',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized"),
  x = c(90, 70, 50, 30, 10, 5),
  textposition = "outside",
  textinfo = "value+percent total") 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized")))

fig
```
### Basic Area Funnel Plot
```r
library(plotly)

fig <- plot_ly(
  type = "funnelarea",
  text = c("The 1st","The 2nd", "The 3rd", "The 4th", "The 5th"),
  values = c(5, 4, 3, 2, 1))

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = "funnelarea",
  text = c("The 1st","The 2nd", "The 3rd", "The 4th", "The 5th"),
  values = c(5, 4, 3, 2, 1))

fig
```
