--- 
title: "Plotly Cheatsheet"
author: "Taichen Zhou, Yichen Huang"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output:
  rmdformats::downcute: 
    self_contained: true
    thumbnails: false
    df_print: paged
    lightbox: true
    gallery: true
    embed_fonts: true
    default_style: "dark"
    toc_depth: 4
    highlight: tango
    code_folding: "show"
    code_download: true
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```

```{css echo=FALSE}
@import url('https://fonts.googleapis.com/css2?family=Fira+Code:wght@300;400;500;600;700&display=swap');
code{
  font-family: 'Fira Code'!important;
}
```

# Introduction

## Getting Start
**plotly** is an R package for creating interactive web-based graphs via the open source JavaScript graphing library plotly.js.
As of version 2.0 (November 17, 2015), graphs created with the plotly R package are rendered locally through the htmlwidgets framework.

## Installation
### CRAN
Use the install.package() function to install the plotly R package from CRAN. This version may not be the absolute latest version, so we recommend downloading from Github using the instructions below if you can.
```
install.packages('plotly')
```
### GitHub
Alternatively, you can install the latest development version of plotly from GitHub via the devtools R package:
```
devtools::install_github("ropensci/plotly")
```

<!--chapter:end:index.Rmd-->

---
output:
  html_document: default
  pdf_document: default
---
```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Basic Charts
## Scatter Plots
### Basic Scatter Plot

```r
library(plotly)

fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg, type = "scatter", mode = "markers")
fig
```

```{r}
library(plotly)

fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg, type = "scatter", mode = "markers")
fig
```
### Styled Scatter Plots

```r
library(plotly)

fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg, type = "scatter",
               mode = "markers"，
               marker = list(size = 10,
                             color = 'rgba(255, 182, 193, .9)',
                             line = list(color = 'rgba(152, 0, 0, .8)',
                                         width = 2)))
fig <- fig %>% layout(title = 'Styled Scatter',
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))

fig
```

```{r}
library(plotly)

fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg, type = "scatter",
               marker = list(size = 10,
                             color = 'rgba(255, 182, 193, .9)',
                             line = list(color = 'rgba(152, 0, 0, .8)',
                                         width = 2)))
fig <- fig %>% layout(title = 'Styled Scatter',
         yaxis = list(zeroline = FALSE),
         xaxis = list(zeroline = FALSE))

fig
```
### Qualitative Colorscales
```r
library(plotly)
fig <- plot_ly(data = mtcars, x = ~hp, y = ~mpg, color = ~cyl)
fig
```

```{r}
library(plotly)
fig <- plot_ly(data = mtcars, x = ~hp, y = ~mpg, color = ~cyl)
fig

```
## Line Plots
### Basic Line Plot
```r
library(plotly)

x <- c(1:100)
random_y <- rnorm(100, mean = 0)
data <- data.frame(x, random_y)

fig <- plot_ly(data, x = ~x, y = ~random_y, type = 'scatter', mode = 'lines')

fig
```
```{r}
library(plotly)

x <- c(1:100)
random_y <- rnorm(100, mean = 0)
data <- data.frame(x, random_y)

fig <- plot_ly(data, x = ~x, y = ~random_y, type = 'scatter', mode = 'lines')

fig
```
### Line Plots Mode
```r
library(plotly)

trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x, y = ~trace_0, name = 'trace 0', type = 'scatter', mode = 'lines') 
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers') 
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')

fig

```
```{r}
library(plotly)

trace_0 <- rnorm(100, mean = 5)
trace_1 <- rnorm(100, mean = 0)
trace_2 <- rnorm(100, mean = -5)
x <- c(1:100)

data <- data.frame(x, trace_0, trace_1, trace_2)

fig <- plot_ly(data, x = ~x, y = ~trace_0, name = 'trace 0', type = 'scatter', mode = 'lines') 
fig <- fig %>% add_trace(y = ~trace_1, name = 'trace 1', mode = 'lines+markers') 
fig <- fig %>% add_trace(y = ~trace_2, name = 'trace 2', mode = 'markers')

fig


```
### Density Plot
```r
library(plotly)

dens <- with(diamonds, tapply(price, INDEX = cut, density))
df <- data.frame(
  x = unlist(lapply(dens, "[[", "x")),
  y = unlist(lapply(dens, "[[", "y")),
  cut = rep(names(dens), each = length(dens[[1]]$x))
)

fig <- plot_ly(df, x = ~x, y = ~y, color = ~cut) 
fig <- fig %>% add_lines()

fig

```

```{r}
library(plotly)

dens <- with(diamonds, tapply(price, INDEX = cut, density))
df <- data.frame(
  x = unlist(lapply(dens, "[[", "x")),
  y = unlist(lapply(dens, "[[", "y")),
  cut = rep(names(dens), each = length(dens[[1]]$x))
)

fig <- plot_ly(df, x = ~x, y = ~y, color = ~cut) 
fig <- fig %>% add_lines()

fig


```

## Bar Charts
### Basic Bar Chart
```r
library(plotly)

fig <- plot_ly(
  x = c("giraffes", "orangutans", "monkeys"),
  y = c(20, 14, 23),
  name = "SF Zoo",
  type = "bar"
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  x = c("giraffes", "orangutans", "monkeys"),
  y = c(20, 14, 23),
  name = "SF Zoo",
  type = "bar"
)

fig

```
### Grouped Bar Chart
```r
library(plotly)

Animals <- c("giraffes", "orangutans", "monkeys")
SF_Zoo <- c(20, 14, 23)
LA_Zoo <- c(12, 18, 29)
data <- data.frame(Animals, SF_Zoo, LA_Zoo)

fig <- plot_ly(data, x = ~Animals, y = ~SF_Zoo, type = 'bar', name = 'SF Zoo')
fig <- fig %>% add_trace(y = ~LA_Zoo, name = 'LA Zoo')
fig <- fig %>% layout(yaxis = list(title = 'Count'), barmode = 'group')

fig
```
```{r}
library(plotly)

Animals <- c("giraffes", "orangutans", "monkeys")
SF_Zoo <- c(20, 14, 23)
LA_Zoo <- c(12, 18, 29)
data <- data.frame(Animals, SF_Zoo, LA_Zoo)

fig <- plot_ly(data, x = ~Animals, y = ~SF_Zoo, type = 'bar', name = 'SF Zoo')
fig <- fig %>% add_trace(y = ~LA_Zoo, name = 'LA Zoo')
fig <- fig %>% layout(yaxis = list(title = 'Count'), barmode = 'group')

fig
```
### Stacked Bar Chart
```r
library(plotly)

Animals <- c("giraffes", "orangutans", "monkeys")
SF_Zoo <- c(20, 14, 23)
LA_Zoo <- c(12, 18, 29)
data <- data.frame(Animals, SF_Zoo, LA_Zoo)

fig <- plot_ly(data, x = ~Animals, y = ~SF_Zoo, type = 'bar', name = 'SF Zoo')
fig <- fig %>% add_trace(y = ~LA_Zoo, name = 'LA Zoo')
fig <- fig %>% layout(yaxis = list(title = 'Count'), barmode = 'stack')

fig
```
```{r}
library(plotly)

Animals <- c("giraffes", "orangutans", "monkeys")
SF_Zoo <- c(20, 14, 23)
LA_Zoo <- c(12, 18, 29)
data <- data.frame(Animals, SF_Zoo, LA_Zoo)

fig <- plot_ly(data, x = ~Animals, y = ~SF_Zoo, type = 'bar', name = 'SF Zoo')
fig <- fig %>% add_trace(y = ~LA_Zoo, name = 'LA Zoo')
fig <- fig %>% layout(yaxis = list(title = 'Count'), barmode = 'stack')

fig
```
### Colored and Styled Bar Chart
```r
library(plotly)

x <- c(1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012)
roW <- c(219, 146, 112, 127, 124, 180, 236, 207, 236, 263, 350, 430, 474, 526, 488, 537, 500, 439)
China <- c(16, 13, 10, 11, 28, 37, 43, 55, 56, 88, 105, 156, 270, 299, 340, 403, 549, 499)
data <- data.frame(x, roW, China)

fig <- plot_ly(data, x = ~x, y = ~roW, type = 'bar', name = 'Rest of the World',
        marker = list(color = 'rgb(55, 83, 109)'))
fig <- fig %>% add_trace(y = ~China, name = 'China', marker = list(color = 'rgb(26, 118, 255)'))
fig <- fig %>% layout(title = 'US Export of Plastic Scrap',
         xaxis = list(
           title = "",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'USD (millions)',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(x = 0, y = 1, bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)

fig
```
```{r}
library(plotly)

x <- c(1995, 1996, 1997, 1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012)
roW <- c(219, 146, 112, 127, 124, 180, 236, 207, 236, 263, 350, 430, 474, 526, 488, 537, 500, 439)
China <- c(16, 13, 10, 11, 28, 37, 43, 55, 56, 88, 105, 156, 270, 299, 340, 403, 549, 499)
data <- data.frame(x, roW, China)

fig <- plot_ly(data, x = ~x, y = ~roW, type = 'bar', name = 'Rest of the World',
        marker = list(color = 'rgb(55, 83, 109)'))
fig <- fig %>% add_trace(y = ~China, name = 'China', marker = list(color = 'rgb(26, 118, 255)'))
fig <- fig %>% layout(title = 'US Export of Plastic Scrap',
         xaxis = list(
           title = "",
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         yaxis = list(
           title = 'USD (millions)',
           titlefont = list(
             size = 16,
             color = 'rgb(107, 107, 107)'),
           tickfont = list(
             size = 14,
             color = 'rgb(107, 107, 107)')),
         legend = list(x = 0, y = 1, bgcolor = 'rgba(255, 255, 255, 0)', bordercolor = 'rgba(255, 255, 255, 0)'),
         barmode = 'group', bargap = 0.15, bargroupgap = 0.1)

fig
```
## Pie Charts
### Basic Pie Chart
```r
library(plotly)

USPersonalExpenditure <- data.frame("Categorie"=rownames(USPersonalExpenditure), USPersonalExpenditure)
data <- USPersonalExpenditure[,c('Categorie', 'X1960')]

fig <- plot_ly(data, labels = ~Categorie, values = ~X1960, type = 'pie')
fig <- fig %>% layout(title = 'United States Personal Expenditures by Categories in 1960',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig

```
```{r}
library(plotly)

USPersonalExpenditure <- data.frame("Categorie"=rownames(USPersonalExpenditure), USPersonalExpenditure)
data <- USPersonalExpenditure[,c('Categorie', 'X1960')]

fig <- plot_ly(data, labels = ~Categorie, values = ~X1960, type = 'pie')
fig <- fig %>% layout(title = 'United States Personal Expenditures by Categories in 1960',
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```
### Donut Chart
```r
library(plotly)
library(dplyr)

# Get Manufacturer
mtcars$manuf <- sapply(strsplit(rownames(mtcars), " "), "[[", 1)

df <- mtcars
df <- df %>% group_by(manuf)
df <- df %>% summarize(count = n())
fig <- df %>% plot_ly(labels = ~manuf, values = ~count)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "Donut charts using Plotly",  showlegend = F,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```
```{r}
library(plotly)
library(dplyr)

# Get Manufacturer
mtcars$manuf <- sapply(strsplit(rownames(mtcars), " "), "[[", 1)

df <- mtcars
df <- df %>% group_by(manuf)
df <- df %>% summarize(count = n())
fig <- df %>% plot_ly(labels = ~manuf, values = ~count)
fig <- fig %>% add_pie(hole = 0.6)
fig <- fig %>% layout(title = "Donut charts using Plotly",  showlegend = F,
                      xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
                      yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))

fig
```
## Bubble Charts
### Simple Bubble Chart
```r
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers',
        marker = list(size = ~Gap, opacity = 0.5))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE))

fig
```
```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers',
        marker = list(size = ~Gap, opacity = 0.5))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE))

fig
```
### Mapping a Color Variable (Continuous)
```r
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers', color = ~Gap, colors = 'Reds',
        marker = list(size = ~Gap, opacity = 0.5))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE))

fig
```
```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers', color = ~Gap, colors = 'Reds',
        marker = list(size = ~Gap, opacity = 0.5))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE))

fig

```
### Mapping a Color Variable (Categorical)
```r
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

data$State <- as.factor(c('Massachusetts', 'California', 'Massachusetts', 'Pennsylvania', 'New Jersey', 'Illinois', 'Washington DC',
                          'Massachusetts', 'Connecticut', 'New York', 'North Carolina', 'New Hampshire', 'New York', 'Indiana',
                          'New York', 'Michigan', 'Rhode Island', 'California', 'Georgia', 'California', 'California'))

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers', size = ~Gap, color = ~State, colors = 'Paired',
        marker = list(opacity = 0.5, sizemode = 'diameter'))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE),
         showlegend = FALSE)

fig
```
```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/school_earnings.csv")

data$State <- as.factor(c('Massachusetts', 'California', 'Massachusetts', 'Pennsylvania', 'New Jersey', 'Illinois', 'Washington DC',
                          'Massachusetts', 'Connecticut', 'New York', 'North Carolina', 'New Hampshire', 'New York', 'Indiana',
                          'New York', 'Michigan', 'Rhode Island', 'California', 'Georgia', 'California', 'California'))

fig <- plot_ly(data, x = ~Women, y = ~Men, text = ~School, type = 'scatter', mode = 'markers', size = ~Gap, color = ~State, colors = 'Paired',
        marker = list(opacity = 0.5, sizemode = 'diameter'))
fig <- fig %>% layout(title = 'Gender Gap in Earnings per University',
         xaxis = list(showgrid = FALSE),
         yaxis = list(showgrid = FALSE),
         showlegend = FALSE)

fig
```
### Styled Buble Chart
```r
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/gapminderDataFiveYear.csv")

data_2007 <- data[which(data$year == 2007),]
data_2007 <- data_2007[order(data_2007$continent, data_2007$country),]
slope <- 2.666051223553066e-05
data_2007$size <- sqrt(data_2007$pop * slope)
colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

fig <- plot_ly(data_2007, x = ~gdpPercap, y = ~lifeExp, color = ~continent, size = ~size, colors = colors,
        type = 'scatter', mode = 'markers', sizes = c(min(data_2007$size), max(data_2007$size)),
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('Country:', country, '<br>Life Expectancy:', lifeExp, '<br>GDP:', gdpPercap,
                      '<br>Pop.:', pop))
fig <- fig %>% layout(title = 'Life Expectancy v. Per Capita GDP, 2007',
         xaxis = list(title = 'GDP per capita (2000 dollars)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(2.003297660701705, 5.191505530708712),
                      type = 'log',
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwidth = 2),
         yaxis = list(title = 'Life Expectancy (years)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(36.12621671352166, 91.72921793264332),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')

fig
```
```{r}
library(plotly)

data <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/gapminderDataFiveYear.csv")

data_2007 <- data[which(data$year == 2007),]
data_2007 <- data_2007[order(data_2007$continent, data_2007$country),]
slope <- 2.666051223553066e-05
data_2007$size <- sqrt(data_2007$pop * slope)
colors <- c('#4AC6B7', '#1972A4', '#965F8A', '#FF7070', '#C61951')

fig <- plot_ly(data_2007, x = ~gdpPercap, y = ~lifeExp, color = ~continent, size = ~size, colors = colors,
        type = 'scatter', mode = 'markers', sizes = c(min(data_2007$size), max(data_2007$size)),
        marker = list(symbol = 'circle', sizemode = 'diameter',
                      line = list(width = 2, color = '#FFFFFF')),
        text = ~paste('Country:', country, '<br>Life Expectancy:', lifeExp, '<br>GDP:', gdpPercap,
                      '<br>Pop.:', pop))
fig <- fig %>% layout(title = 'Life Expectancy v. Per Capita GDP, 2007',
         xaxis = list(title = 'GDP per capita (2000 dollars)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(2.003297660701705, 5.191505530708712),
                      type = 'log',
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwidth = 2),
         yaxis = list(title = 'Life Expectancy (years)',
                      gridcolor = 'rgb(255, 255, 255)',
                      range = c(36.12621671352166, 91.72921793264332),
                      zerolinewidth = 1,
                      ticklen = 5,
                      gridwith = 2),
         paper_bgcolor = 'rgb(243, 243, 243)',
         plot_bgcolor = 'rgb(243, 243, 243)')

fig
```
## Sankey Diagram
### Basic Sankey Diagram
```r
library(plotly)

fig <- plot_ly(
    type = "sankey",
    orientation = "h",

    node = list(
      label = c("A1", "A2", "B1", "B2", "C1", "C2"),
      color = c("blue", "blue", "blue", "blue", "blue", "blue"),
      pad = 15,
      thickness = 20,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = c(0,1,0,2,3,3),
      target = c(2,3,3,4,4,5),
      value =  c(8,4,2,8,4,2)
    )
  )
fig <- fig %>% layout(
    title = "Basic Sankey Diagram",
    font = list(
      size = 10
    )
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = "sankey",
    orientation = "h",

    node = list(
      label = c("A1", "A2", "B1", "B2", "C1", "C2"),
      color = c("blue", "blue", "blue", "blue", "blue", "blue"),
      pad = 15,
      thickness = 20,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = c(0,1,0,2,3,3),
      target = c(2,3,3,4,4,5),
      value =  c(8,4,2,8,4,2)
    )
  )
fig <- fig %>% layout(
    title = "Basic Sankey Diagram",
    font = list(
      size = 10
    )
)

fig
```
### Style Sankey Diagram
```r
library(plotly)
library(rjson)

json_file <- "https://raw.githubusercontent.com/plotly/plotly.js/master/test/image/mocks/sankey_energy_dark.json"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

fig <- plot_ly(
    type = "sankey",
    domain = list(
      x =  c(0,1),
      y =  c(0,1)
    ),
    orientation = "h",
    valueformat = ".0f",
    valuesuffix = "TWh",

    node = list(
      label = json_data$data[[1]]$node$label,
      color = json_data$data[[1]]$node$color,
      pad = 15,
      thickness = 15,
      line = list(
        color = "black",
        width = 0.5
      )
    ),

    link = list(
      source = json_data$data[[1]]$link$source,
      target = json_data$data[[1]]$link$target,
      value =  json_data$data[[1]]$link$value,
      label =  json_data$data[[1]]$link$label
    )
  )
fig <- fig %>% layout(
    title = "Energy forecast for 2050<br>Source: Department of Energy & Climate Change, Tom Counsell via <a href='https://bost.ocks.org/mike/sankey/'>Mike Bostock</a>",
    font = list(
      size = 10,
      color = 'white'
    ),
    xaxis = list(showgrid = F, zeroline = F, showticklabels = F),
    yaxis = list(showgrid = F, zeroline = F, showticklabels = F),
    plot_bgcolor = 'black',
    paper_bgcolor = 'black'
)

fig
```
```{r}
library(plotly)
library(rjson)

json_file <- "https://raw.githubusercontent.com/plotly/plotly.js/master/test/image/mocks/sankey_energy_dark.json"
json_data <- fromJSON(paste(readLines(json_file), collapse=""))

fig <- plot_ly(
    type = "sankey",
    domain = list(
      x =  c(0,1),
      y =  c(0,1)
    ),
    orientation = "h",
    valueformat = ".0f",
    valuesuffix = "TWh",

    node = list(
      label = json_data$data[[1]]$node$label,
      color = json_data$data[[1]]$node$color,
      pad = 15,
      thickness = 15,
      line = list(
        color = "white",
        width = 0.5
      )
    ),

    link = list(
      source = json_data$data[[1]]$link$source,
      target = json_data$data[[1]]$link$target,
      value =  json_data$data[[1]]$link$value,
      label =  json_data$data[[1]]$link$label
    )
  )
fig <- fig %>% layout(
    title = "Energy forecast for 2050<br>Source: Department of Energy & Climate Change, Tom Counsell via <a href='https://bost.ocks.org/mike/sankey/'>Mike Bostock</a>",
    font = list(
      size = 10,
      color = 'white'
    ),
    xaxis = list(showgrid = F, zeroline = F, showticklabels = F),
    yaxis = list(showgrid = F, zeroline = F, showticklabels = F),
    plot_bgcolor = 'black',
    paper_bgcolor = 'black'
)

fig
```



<!--chapter:end:basic.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Statistical Charts

## 2D Histograms

A 2D histogram is a visualization of a bivariate distribution. And it looks like a heat map.

``` r
#
library(plotly)

s <- matrix(c(1, -.75, -.75, 1), ncol = 2)
obs <- mvtnorm::rmvnorm(500, sigma = s)
fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg)
fig2 <- subplot(
  fig %>% add_markers(alpha = 0.2),
  fig %>% add_histogram2d()
)

fig2
```

```{r}
library(plotly)

fig <- plot_ly(data = mtcars, x = ~hp, y= ~mpg)
fig2 <- subplot(
  fig %>% add_markers(alpha = 0.5),
  fig %>% add_histogram2d(colorscale = "Blues")
)

fig2

```

## Box Plot

### Basic Boxplot

``` r
library(plotly)
fig <- plot_ly(data = mtcars, y= ~mpg, type = "box")

fig
```

```{r}
library(plotly)
fig <- plot_ly(data = mtcars, y= ~mpg, type = "box")

fig
```

### Horizontal Boxplot

``` r
library(plotly)
fig <- plot_ly(data = mtcars, x= ~mpg, type = "box")

fig
```

```{r}
library(plotly)
fig <- plot_ly(data = mtcars, x= ~mpg, type = "box")

fig
```

### Adding Jittered Points

``` r
library(plotly)
fig <- plot_ly(data = mtcars, y= ~mpg, type = "box", boxpoints = "all", jitter = 0.3)

fig
```

```{r}
library(plotly)
fig <- plot_ly(data = mtcars, y= ~mpg, type = "box", boxpoints = "all", jitter = 0.3)

fig

```

### Several Box Plots

``` r
fig <- plot_ly(data = mtcars, y= ~mpg, color = ~as.factor(cyl), type = "box")

fig
```

```{r}
fig <- plot_ly(data = mtcars, y= ~mpg, color = ~as.factor(cyl), type = "box")

fig
```

## Histograms

### Basic Histograms

``` r
library(plotly)
fig <- plot_ly(data = mtcars, x = ~hp, type = "histogram", nbinsx = 15)

fig
```

```{r}
library(plotly)
fig <- plot_ly(data = mtcars, x = ~hp, type = "histogram", nbinsx = 15)

fig

```

### Stacked Histogram

``` r

fig <- plot_ly(data = mtcars, x = ~hp,
             type = "histogram",
             cumulative = list(enabled=TRUE),
             nbinsx = 15)

fig
```

```{r}

fig <- plot_ly(data = mtcars, x = ~hp,
             type = "histogram",
             cumulative = list(enabled=TRUE),
             nbinsx = 15)

fig
```

## Violin Plots

``` r
fig <- mtcars %>%
  plot_ly(
    y = ~hp,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    x0 = 'Horse Power'
  ) 

fig <- fig %>%
  layout(
    yaxis = list(
      title = "",
      zeroline = F
    )
  )

fig
```

```{r}
fig <- mtcars %>%
  plot_ly(
    y = ~hp,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    ),
    x0 = 'Horse Power'
  ) 

fig <- fig %>%
  layout(
    yaxis = list(
      title = "",
      zeroline = F
    )
  )

fig

```

<!--chapter:end:statistical.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Scientific Charts
## Log Plots
### Log Axes
```r
library(plotly)
d <- diamonds[sample(nrow(diamonds), 1000), ]
# without log scales
fig <- plot_ly(d, x = ~carat, y = ~price) %>% add_markers()

fig
```
```{r}
library(plotly)
d <- diamonds[sample(nrow(diamonds), 1000), ]
# without log scales
fig <- plot_ly(d, x = ~carat, y = ~price) %>% add_markers()

fig
```
```r
# with log scales
fig <- layout(fig, xaxis = list(type = "log"),
       yaxis = list(type = "log"))

fig
```
```{r}
# with log scales
fig <- layout(fig, xaxis = list(type = "log"),
       yaxis = list(type = "log"))

fig
```

## Contour Plots
### Basic Contour
```r
library(plotly)

fig <- plot_ly(z = ~volcano, type = "contour")

fig
```
```{r}
library(plotly)

fig <- plot_ly(z = ~volcano, type = "contour")

fig
```
### Smoothing Contour Coloring
```r
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z = matrix(c(10, 10.625, 12.5, 15.625, 20, 5.625, 6.25, 8.125, 
               11.25, 15.625, 2.5, 3.125, 5, 8.125, 12.5, 0.625, 
               1.25, 3.125, 6.25, 10.625, 0, 0.625, 2.5, 5.625, 
               10), nrow=5, ncol=5),
  contours = list(
    coloring = 'heatmap'
  )
)

fig

```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z = matrix(c(10, 10.625, 12.5, 15.625, 20, 5.625, 6.25, 8.125, 
               11.25, 15.625, 2.5, 3.125, 5, 8.125, 12.5, 0.625, 
               1.25, 3.125, 6.25, 10.625, 0, 0.625, 2.5, 5.625, 
               10), nrow=5, ncol=5),
  contours = list(
    coloring = 'heatmap'
  )
)

fig

```

## Heatmaps
### Basic Heatmap
```r
library(plotly)
fig <- plot_ly(z = volcano, type = "heatmap")

fig
```
```{r}
library(plotly)
fig <- plot_ly(z = volcano, type = "heatmap")

fig
```
### Sequential Colorscales: Greys
```r
fig <- plot_ly(z = volcano, colors = "Greys", type = "heatmap")

fig
```
```{r}
fig <- plot_ly(z = volcano, colors = "Greys", type = "heatmap")

fig
```

## Parallel Coordinates
### Basic Parallel Cordinates Plots
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/iris.csv")

fig <- df %>% plot_ly(type = 'parcoords',
          line = list(color = ~species_id,
                      colorscale = list(c(0,'red'),c(0.5,'green'),c(1,'blue'))),
          dimensions = list(
            list(range = c(2,4.5),
                 label = 'Sepal Width', values = ~sepal_width),
            list(range = c(4,8),
                 constraintrange = c(5,6),
                 label = 'Sepal Length', values = ~sepal_length),
            list(range = c(0,2.5),
                 label = 'Petal Width', values = ~petal_width),
            list(range = c(1,7),
                 label = 'Petal Length', values = ~petal_length)
            )
          )

fig
```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/iris.csv")

fig <- df %>% plot_ly(type = 'parcoords',
          line = list(color = ~species_id,
                      colorscale = list(c(0,'red'),c(0.5,'green'),c(1,'blue'))),
          dimensions = list(
            list(range = c(2,4.5),
                 label = 'Sepal Width', values = ~sepal_width),
            list(range = c(4,8),
                 constraintrange = c(5,6),
                 label = 'Sepal Length', values = ~sepal_length),
            list(range = c(0,2.5),
                 label = 'Petal Width', values = ~petal_width),
            list(range = c(1,7),
                 label = 'Petal Length', values = ~petal_length)
            )
          )

fig
```
### Advanced Parallel Coordinates Plot
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/parcoords_data.csv")

fig <- df %>%
  plot_ly(width = 1000, height = 600) 
fig <- fig %>% add_trace(type = 'parcoords',
          line = list(color = ~colorVal,
                      colorscale = 'Jet',
                      showscale = TRUE,
                      reversescale = TRUE,
                      cmin = -4000,
                      cmax = -100),
          dimensions = list(
            list(range = c(~min(blockHeight),~max(blockHeight)),
                 constraintrange = c(100000,150000),
                 label = 'Block Height', values = ~blockHeight),
            list(range = c(~min(blockWidth),~max(blockWidth)),
                 label = 'Block Width', values = ~blockWidth),
            list(tickvals = c(0,0.5,1,2,3),
                 ticktext = c('A','AB','B','Y','Z'),
                 label = 'Cyclinder Material', values = ~cycMaterial),
            list(range = c(-1,4),
                 tickvals = c(0,1,2,3),
                 label = 'Block Material', values = ~blockMaterial),
            list(range = c(~min(totalWeight),~max(totalWeight)),
                 visible = TRUE,
                 label = 'Total Weight', values = ~totalWeight),
            list(range = c(~min(assemblyPW),~max(assemblyPW)),
                 label = 'Assembly Penalty Weight', values = ~assemblyPW),
            list(range = c(~min(HstW),~max(HstW)),
                 label = 'Height st Width', values = ~HstW),
            list(range = c(~min(minHW),~max(minHW)),
                 label = 'Min Height Width', values = ~minHW),
            list(range = c(~min(minWD),~max(minWD)),
                 label = 'Min Width Diameter', values = ~minWD),
            list(range = c(~min(rfBlock),~max(rfBlock)),
                 label = 'RF Block', values = ~rfBlock)
            )
          )


fig

```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/parcoords_data.csv")

fig <- df %>%
  plot_ly(width = 1000, height = 600) 
fig <- fig %>% add_trace(type = 'parcoords',
          line = list(color = ~colorVal,
                      colorscale = 'Jet',
                      showscale = TRUE,
                      reversescale = TRUE,
                      cmin = -4000,
                      cmax = -100),
          dimensions = list(
            list(range = c(~min(blockHeight),~max(blockHeight)),
                 constraintrange = c(100000,150000),
                 label = 'Block Height', values = ~blockHeight),
            list(range = c(~min(blockWidth),~max(blockWidth)),
                 label = 'Block Width', values = ~blockWidth),
            list(tickvals = c(0,0.5,1,2,3),
                 ticktext = c('A','AB','B','Y','Z'),
                 label = 'Cyclinder Material', values = ~cycMaterial),
            list(range = c(-1,4),
                 tickvals = c(0,1,2,3),
                 label = 'Block Material', values = ~blockMaterial),
            list(range = c(~min(totalWeight),~max(totalWeight)),
                 visible = TRUE,
                 label = 'Total Weight', values = ~totalWeight),
            list(range = c(~min(assemblyPW),~max(assemblyPW)),
                 label = 'Assembly Penalty Weight', values = ~assemblyPW),
            list(range = c(~min(HstW),~max(HstW)),
                 label = 'Height st Width', values = ~HstW),
            list(range = c(~min(minHW),~max(minHW)),
                 label = 'Min Height Width', values = ~minHW),
            list(range = c(~min(minWD),~max(minWD)),
                 label = 'Min Width Diameter', values = ~minWD),
            list(range = c(~min(rfBlock),~max(rfBlock)),
                 label = 'RF Block', values = ~rfBlock)
            )
          )


fig

```
## Polar Charts
### Basic Polar Charts
```r
library(plotly)

fig <- plot_ly(
  type = 'scatterpolar',
  r = c(0,1,2,2),
  theta = c(0,45,90,0),
  mode = 'markers'
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatterpolar',
  r = c(0,1,2,2),
  theta = c(0,45,90,0),
  mode = 'markers'
)

fig
```
### Line Polar Charts
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/polar_dataset.csv")

fig <- plot_ly(
  df,
  type = 'scatterpolar',
  mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = ~x1,
    theta = ~y,
    name = 'Figure8',
    line = list(
      color = 'peru'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x2,
    theta = ~y,
    name = 'Cardioid',
    line = list(
      color = 'darkviolet'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x3,
    theta = ~y,
    name = 'Hypercardioid',
    line = list(
      color = 'deepskyblue'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x4,
    theta = ~y,
    name = 'Subcardioid',
    line = list(
      color = 'orangered'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x5,
    theta = ~y,
    name = 'Supercardioid',
    line = list(
      color = 'green'
    )
  ) 

fig <- fig %>%
  layout(
    title = 'Mic Patterns',
    font = list(
      family = 'Arial',
      size = 12,
      color = '#000'
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/polar_dataset.csv")

fig <- plot_ly(
  df,
  type = 'scatterpolar',
  mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = ~x1,
    theta = ~y,
    name = 'Figure8',
    line = list(
      color = 'peru'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x2,
    theta = ~y,
    name = 'Cardioid',
    line = list(
      color = 'darkviolet'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x3,
    theta = ~y,
    name = 'Hypercardioid',
    line = list(
      color = 'deepskyblue'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x4,
    theta = ~y,
    name = 'Subcardioid',
    line = list(
      color = 'orangered'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x5,
    theta = ~y,
    name = 'Supercardioid',
    line = list(
      color = 'green'
    )
  ) 

fig <- fig %>%
  layout(
    title = 'Mic Patterns',
    font = list(
      family = 'Arial',
      size = 12,
      color = '#000'
    ),
    showlegend = F
  )

fig
```
### Area Polar Area
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 1.5, 1.5, 0, 2.5, 2.5, 0),
    theta = c(0, 10, 25, 0, 205, 215, 0),
    fill = 'toself',
    fillcolor = '#709Bff',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3.5, 3.5, 0),
    theta = c(0, 55, 75, 0),
    fill = 'toself',
    fillcolor = '#E4FF87',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4.5, 4.5, 0, 4.5, 4.5, 0),
    theta = c(0, 100, 120, 0, 305, 320, 0),
    fill = 'toself',
    fillcolor = '#FFAA70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4, 4, 0),
    theta = c(0, 165, 195, 0),
    fill = 'toself',
    fillcolor = '#FFDF70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3, 3, 0),
    theta = c(0, 262.5, 277.5, 0),
    fill = 'toself',
    fillcolor = '#B6FFB4',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,5)
      )
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 1.5, 1.5, 0, 2.5, 2.5, 0),
    theta = c(0, 10, 25, 0, 205, 215, 0),
    fill = 'toself',
    fillcolor = '#709Bff',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3.5, 3.5, 0),
    theta = c(0, 55, 75, 0),
    fill = 'toself',
    fillcolor = '#E4FF87',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4.5, 4.5, 0, 4.5, 4.5, 0),
    theta = c(0, 100, 120, 0, 305, 320, 0),
    fill = 'toself',
    fillcolor = '#FFAA70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4, 4, 0),
    theta = c(0, 165, 195, 0),
    fill = 'toself',
    fillcolor = '#FFDF70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3, 3, 0),
    theta = c(0, 262.5, 277.5, 0),
    fill = 'toself',
    fillcolor = '#B6FFB4',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,5)
      )
    ),
    showlegend = F
  )

fig
```

## Radar Charts
### Basic Radar Charts
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    fill = 'toself'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    fill = 'toself'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    ),
    showlegend = F
  )

fig
```
### Multiple Trace Radar Charts
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    fill = 'toself'
  ) 
fig <- fig %>%
  add_trace(
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group A'
  ) 
fig <- fig %>%
  add_trace(
    r = c(1.5, 10, 39, 31, 15, 1.5),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group B'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    )
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    fill = 'toself'
  ) 
fig <- fig %>%
  add_trace(
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group A'
  ) 
fig <- fig %>%
  add_trace(
    r = c(1.5, 10, 39, 31, 15, 1.5),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group B'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    )
  )

fig
```


<!--chapter:end:scientific.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Financial Charts
## Time Series and Date Axess
### Time Series using Axes of type date
```r
library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', width = 900)


fig
```
```{r}
library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', width = 900)


fig
```
### Configuring Tick Labels
```r
library(tidyquant)
library(plotly)
tickers = c("GOOG", "AAPL", "AMZN", "META", "NFLX", "MSFT")
for (i in tickers){
  getSymbols(i,
             from = "2020-01-01",
             to = "2021-12-31")}
stock <- data.frame(GOOG$GOOG.Adjusted,
                    AAPL$AAPL.Adjusted,
                    AMZN$AMZN.Adjusted,
                    META$META.Adjusted,
                    NFLX$NFLX.Adjusted,
                    MSFT$MSFT.Adjusted)
stock$GOOG.Adjusted <- stock$GOOG.Adjusted/stock$GOOG.Adjusted[1]
stock$AAPL.Adjusted <- stock$AAPL.Adjusted/stock$AAPL.Adjusted[1]
stock$AMZN.Adjusted <- stock$AMZN.Adjusted/stock$AMZN.Adjusted[1]
stock$META.Adjusted <- stock$META.Adjusted/stock$META.Adjusted[1]
stock$NFLX.Adjusted <- stock$NFLX.Adjusted/stock$NFLX.Adjusted[1]
stock$MSFT.Adjusted <- stock$MSFT.Adjusted/stock$MSFT.Adjusted[1]
stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Dates, y = ~GOOG, name = 'GOOG')%>%
  add_trace(x = ~Dates, y = ~AAPL, name = 'AAPL')%>%
  add_trace(x = ~Dates, y = ~AMZN, name = 'AMZN')%>%
  add_trace(x = ~Dates, y = ~META, name = 'META')%>%
  add_trace(x = ~Dates, y = ~NFLX, name = 'NFLX')%>%
  add_trace(x = ~Dates, y = ~MSFT, name = 'MSFT')%>%
  layout(title = 'custom tick labels',legend=list(title=list(text='variable')),
         xaxis = list(dtick = "M1", tickformat="%b<br>%Y"), width = 1000)
options(warn = -1)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6')


fig
```
```{r}
library(tidyquant)
library(plotly)
tickers = c("GOOG", "AAPL", "AMZN", "META", "NFLX", "MSFT")
for (i in tickers){
  getSymbols(i,
             from = "2020-01-01",
             to = "2021-12-31")}
stock <- data.frame(GOOG$GOOG.Adjusted,
                    AAPL$AAPL.Adjusted,
                    AMZN$AMZN.Adjusted,
                    META$META.Adjusted,
                    NFLX$NFLX.Adjusted,
                    MSFT$MSFT.Adjusted)
stock$GOOG.Adjusted <- stock$GOOG.Adjusted/stock$GOOG.Adjusted[1]
stock$AAPL.Adjusted <- stock$AAPL.Adjusted/stock$AAPL.Adjusted[1]
stock$AMZN.Adjusted <- stock$AMZN.Adjusted/stock$AMZN.Adjusted[1]
stock$META.Adjusted <- stock$META.Adjusted/stock$META.Adjusted[1]
stock$NFLX.Adjusted <- stock$NFLX.Adjusted/stock$NFLX.Adjusted[1]
stock$MSFT.Adjusted <- stock$MSFT.Adjusted/stock$MSFT.Adjusted[1]
stock <- data.frame(stock,rownames(stock))
colnames(stock) <- append(tickers,'Dates')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Dates, y = ~GOOG, name = 'GOOG')%>%
  add_trace(x = ~Dates, y = ~AAPL, name = 'AAPL')%>%
  add_trace(x = ~Dates, y = ~AMZN, name = 'AMZN')%>%
  add_trace(x = ~Dates, y = ~META, name = 'META')%>%
  add_trace(x = ~Dates, y = ~NFLX, name = 'NFLX')%>%
  add_trace(x = ~Dates, y = ~MSFT, name = 'MSFT')%>%
  layout(title = 'custom tick labels',legend=list(title=list(text='variable')),
         xaxis = list(dtick = "M1", tickformat="%b<br>%Y"), width = 1000)
options(warn = -1)
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6')


fig
```
### Time Series with Range Selector Buttons
```r

library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F, title='Time Series with Range Slider and Selectors',
         xaxis = list(rangeslider = list(visible = T),
                      rangeselector=list(
                        buttons=list(
                          list(count=1, label="1m", step="month", stepmode="backward"),
                          list(count=6, label="6m", step="month", stepmode="backward"),
                          list(count=1, label="YTD", step="year", stepmode="todate"),
                          list(count=1, label="1y", step="year", stepmode="backward"),
                          list(step="all")
                        ))))
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', margin = 0.1, width = 900)
fig
```
```{r}

library(plotly)

stock <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/finance-charts-apple.csv')

fig <- plot_ly(stock, type = 'scatter', mode = 'lines')%>%
  add_trace(x = ~Date, y = ~AAPL.High)%>%
  layout(showlegend = F, title='Time Series with Range Slider and Selectors',
         xaxis = list(rangeslider = list(visible = T),
                      rangeselector=list(
                        buttons=list(
                          list(count=1, label="1m", step="month", stepmode="backward"),
                          list(count=6, label="6m", step="month", stepmode="backward"),
                          list(count=1, label="YTD", step="year", stepmode="todate"),
                          list(count=1, label="1y", step="year", stepmode="backward"),
                          list(step="all")
                        ))))
fig <- fig %>%
  layout(
         xaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         yaxis = list(zerolinecolor = '#ffff',
                      zerolinewidth = 2,
                      gridcolor = 'ffff'),
         plot_bgcolor='#e5ecf6', margin = 0.1, width = 900)
fig
```

## Candlestick Charts
### Basic Candlestick
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
# basic example of ohlc charts
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic Candlestick Chart")

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
# basic example of ohlc charts
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic Candlestick Chart")

fig
```
### Customise the fig ure with Shapes and Annotations
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
)

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Apple Stock",
         annotations = a,
         shapes = l)

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
)

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
)

fig <- df %>% plot_ly(x = ~Date, type="candlestick",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Apple Stock",
         annotations = a,
         shapes = l)

fig
```

## OHLC Charts
### Basic OHLC Chart
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic OHLC Chart",
         xaxis = list(rangeslider = list(visible = F)))

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))
df <- tail(df, 30)

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Basic OHLC Chart",
         xaxis = list(rangeslider = list(visible = F)))

fig
```
### Customise the Figure with Shapes and Annotations
```r
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
          )

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
          )

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Custom Colors",
         annotations = a,
         shapes = l)

fig
```
```{r}
library(plotly)
library(quantmod)

getSymbols("AAPL",src='yahoo')
df <- data.frame(Date=index(AAPL),coredata(AAPL))

# annotation
a <- list(text = "Stock Split",
          x = '2014-06-06',
          y = 1.02,
          xref = 'x',
          yref = 'paper',
          xanchor = 'left',
          showarrow = FALSE
          )

# use shapes to create a line
l <- list(type = line,
          x0 = '2014-06-06',
          x1 = '2014-06-06',
          y0 = 0,
          y1 = 1,
          xref = 'x',
          yref = 'paper',
          line = list(color = 'black',
                      width = 0.5)
          )

fig <- df %>% plot_ly(x = ~Date, type="ohlc",
          open = ~AAPL.Open, close = ~AAPL.Close,
          high = ~AAPL.High, low = ~AAPL.Low) 
fig <- fig %>% layout(title = "Custom Colors",
         annotations = a,
         shapes = l)

fig
```


## Waterfall Charts
### Basic Waterfall Chart
```r
library(plotly)

x= list("Sales", "Consulting", "Net revenue", "Purchases", "Other expenses", "Profit before tax")
measure= c("relative", "relative", "total", "relative", "relative", "total")
text= c("+60", "+80", "", "-40", "-20", "Total")
y= c(60, 80, 0, -40, -20, 0)
data = data.frame(x=factor(x,levels=x),measure,text,y)

fig <- plot_ly(
  data, name = "20", type = "waterfall", measure = ~measure,
  x = ~x, textposition = "outside", y= ~y, text =~text,
  connector = list(line = list(color= "rgb(63, 63, 63)"))) 
fig <- fig %>%
  layout(title = "Profit and loss statement 2018",
        xaxis = list(title = ""),
        yaxis = list(title = ""),
        autosize = TRUE,
        showlegend = TRUE)

fig
```
```{r}
library(plotly)

x= list("Sales", "Consulting", "Net revenue", "Purchases", "Other expenses", "Profit before tax")
measure= c("relative", "relative", "total", "relative", "relative", "total")
text= c("+60", "+80", "", "-40", "-20", "Total")
y= c(60, 80, 0, -40, -20, 0)
data = data.frame(x=factor(x,levels=x),measure,text,y)

fig <- plot_ly(
  data, name = "20", type = "waterfall", measure = ~measure,
  x = ~x, textposition = "outside", y= ~y, text =~text,
  connector = list(line = list(color= "rgb(63, 63, 63)"))) 
fig <- fig %>%
  layout(title = "Profit and loss statement 2018",
        xaxis = list(title = ""),
        yaxis = list(title = ""),
        autosize = TRUE,
        showlegend = TRUE)

fig
```

## Funnel Charts
```r
# Need to install plotly from Github to get funnel plots
devtools::install_github("ropensci/plotly")
```
### Basic Funnel Plot
```r
library(plotly)

fig <- plot_ly() 
fig <- fig %>%
  add_trace(
  type = "funnel",
  y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
  x = c(39, 27.4, 20.6, 11, 2)) 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent")))

fig
```
```{r}
library(plotly)

fig <- plot_ly() 
fig <- fig %>%
  add_trace(
  type = "funnel",
  y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
  x = c(39, 27.4, 20.6, 11, 2)) 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent")))

fig
```
### Stacked Funnel Plot
```r
library(plotly)

fig <- plot_ly(
    type = "funnel",
    name = 'Montreal',
    y = c("Website visit", "Downloads", "Potential customers", "Requested price"),
    x = c(120, 60, 30, 20),
    textinfo = "value+percent initial") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Toronto',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
    x = c(100, 60, 40, 30, 20),
    textposition = "inside",
    textinfo = "value+percent previous") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Vancouver',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized"),
  x = c(90, 70, 50, 30, 10, 5),
  textposition = "outside",
  textinfo = "value+percent total") 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized")))

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = "funnel",
    name = 'Montreal',
    y = c("Website visit", "Downloads", "Potential customers", "Requested price"),
    x = c(120, 60, 30, 20),
    textinfo = "value+percent initial") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Toronto',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent"),
    x = c(100, 60, 40, 30, 20),
    textposition = "inside",
    textinfo = "value+percent previous") 
fig <- fig %>%
  add_trace(
    type = "funnel",
    name = 'Vancouver',
    orientation = "h",
    y = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized"),
  x = c(90, 70, 50, 30, 10, 5),
  textposition = "outside",
  textinfo = "value+percent total") 
fig <- fig %>%
  layout(yaxis = list(categoryarray = c("Website visit", "Downloads", "Potential customers", "Requested price", "invoice sent", "Finalized")))

fig
```
### Basic Area Funnel Plot
```r
library(plotly)

fig <- plot_ly(
  type = "funnelarea",
  text = c("The 1st","The 2nd", "The 3rd", "The 4th", "The 5th"),
  values = c(5, 4, 3, 2, 1))

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = "funnelarea",
  text = c("The 1st","The 2nd", "The 3rd", "The 4th", "The 5th"),
  values = c(5, 4, 3, 2, 1))

fig
```

<!--chapter:end:financial.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Maps
## Choropleth Maps
### Choropleth Map Using GeoJSON
```r
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
df <- read.csv(url2, colClasses=c(fips="character"))
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_ly()
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=counties,
    locations=df$fips,
    z=df$unemp,
    colorscale="Viridis",
    zmin=0,
    zmax=12,
    marker=list(line=list(
      width=0)
    )
  )
fig <- fig %>% colorbar(title = "Unemployment Rate (%)")
fig <- fig %>% layout(
    title = "2016 US Unemployment by County"
)

fig <- fig %>% layout(
    geo = g
  )

fig

```
```{r}
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/geojson-counties-fips.json'
counties <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/fips-unemp-16.csv"
df <- read.csv(url2, colClasses=c(fips="character"))
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)
fig <- plot_ly()
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=counties,
    locations=df$fips,
    z=df$unemp,
    colorscale="Viridis",
    zmin=0,
    zmax=12,
    marker=list(line=list(
      width=0)
    )
  )
fig <- fig %>% colorbar(title = "Unemployment Rate (%)")
fig <- fig %>% layout(
    title = "2016 US Unemployment by County"
)

fig <- fig %>% layout(
    geo = g
  )

fig

```
### Indexing by GeoJSON Properties
```r
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/election.geojson'
geojson <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/election.csv"
df <- read.csv(url2)
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig <- plot_ly() 
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=geojson,
    locations=df$district,
    z=df$Bergeron,
    colorscale="Viridis",
    featureidkey="properties.district"
  )
fig <- fig %>% layout(
    geo = g
  )
fig <- fig %>% colorbar(title = "Bergeron Votes")
fig <- fig %>% layout(
    title = "2013 Montreal Election"
)
fig
```
```{r}
library(plotly)
library(rjson)

url <- 'https://raw.githubusercontent.com/plotly/datasets/master/election.geojson'
geojson <- rjson::fromJSON(file=url)
url2<- "https://raw.githubusercontent.com/plotly/datasets/master/election.csv"
df <- read.csv(url2)
g <- list(
  fitbounds = "locations",
  visible = FALSE
)
fig <- plot_ly() 
fig <- fig %>% add_trace(
    type="choropleth",
    geojson=geojson,
    locations=df$district,
    z=df$Bergeron,
    colorscale="Viridis",
    featureidkey="properties.district"
  )
fig <- fig %>% layout(
    geo = g
  )
fig <- fig %>% colorbar(title = "Bergeron Votes")
fig <- fig %>% layout(
    title = "2013 Montreal Election"
)
fig
```
### Customize choropleth chart
```r
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Millions USD")
fig <- fig %>% layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)',
    geo = g
  )

fig
```
```{r}
library(plotly)
df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/2011_us_ag_exports.csv")
df$hover <- with(df, paste(state, '<br>', "Beef", beef, "Dairy", dairy, "<br>",
                           "Fruits", total.fruits, "Veggies", total.veggies,
                           "<br>", "Wheat", wheat, "Corn", corn))
# give state boundaries a white border
l <- list(color = toRGB("white"), width = 2)
# specify some map projection/options
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showlakes = TRUE,
  lakecolor = toRGB('white')
)

fig <- plot_geo(df, locationmode = 'USA-states')
fig <- fig %>% add_trace(
    z = ~total.exports, text = ~hover, locations = ~code,
    color = ~total.exports, colors = 'Purples'
  )
fig <- fig %>% colorbar(title = "Millions USD")
fig <- fig %>% layout(
    title = '2011 US Agriculture Exports by State<br>(Hover for breakdown)',
    geo = g
  )

fig
```
### World Choropleth Map
```r

df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
fig <- fig %>% add_trace(
    z = ~GDP..BILLIONS., color = ~GDP..BILLIONS., colors = 'Blues',
    text = ~COUNTRY, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'GDP Billions US$', tickprefix = '$')
fig <- fig %>% layout(
    title = '2014 Global GDP<br>Source:<a href="https://www.cia.gov/library/publications/the-world-factbook/fields/2195.html">CIA World Factbook</a>',
    geo = g
  )

fig
```
```{r}

df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_world_gdp_with_codes.csv')

# light grey boundaries
l <- list(color = toRGB("grey"), width = 0.5)

# specify map projection/options
g <- list(
  showframe = FALSE,
  showcoastlines = FALSE,
  projection = list(type = 'Mercator')
)

fig <- plot_geo(df)
fig <- fig %>% add_trace(
    z = ~GDP..BILLIONS., color = ~GDP..BILLIONS., colors = 'Blues',
    text = ~COUNTRY, locations = ~CODE, marker = list(line = l)
  )
fig <- fig %>% colorbar(title = 'GDP Billions US$', tickprefix = '$')
fig <- fig %>% layout(
    title = '2014 Global GDP<br>Source:<a href="https://www.cia.gov/library/publications/the-world-factbook/fields/2195.html">CIA World Factbook</a>',
    geo = g
  )

fig
```

## Scatter Plots on Maps
### Basic Scatter on Map
```r
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2011_february_us_airport_traffic.csv')

# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

fig <- plot_geo(df, lat = ~lat, lon = ~long)
fig <- fig %>% add_markers(
    text = ~paste(airport, city, state, paste("Arrivals:", cnt), sep = "<br />"),
    color = ~cnt, symbol = I("square"), size = I(8), hoverinfo = "text"
  )
fig <- fig %>% colorbar(title = "Incoming flights<br />February 2011")
fig <- fig %>% layout(
    title = 'Most trafficked US airports<br />(Hover for airport)', geo = g
  )

fig
```
```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2011_february_us_airport_traffic.csv')

# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

fig <- plot_geo(df, lat = ~lat, lon = ~long)
fig <- fig %>% add_markers(
    text = ~paste(airport, city, state, paste("Arrivals:", cnt), sep = "<br />"),
    color = ~cnt, symbol = I("square"), size = I(8), hoverinfo = "text"
  )
fig <- fig %>% colorbar(title = "Incoming flights<br />February 2011")
fig <- fig %>% layout(
    title = 'Most trafficked US airports<br />(Hover for airport)', geo = g
  )

fig
```
### Style Scatter Map Layout
```r
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2015_06_30_precipitation.csv')

# change default color scale title
m <- list(colorbar = list(title = "Total Inches"))

# geo styling
g <- list(
  scope = 'north america',
  showland = TRUE,
  landcolor = toRGB("grey83"),
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white"),
  showlakes = TRUE,
  lakecolor = toRGB("white"),
  showsubunits = TRUE,
  showcountries = TRUE,
  resolution = 50,
  projection = list(
    type = 'conic conformal',
    rotation = list(lon = -100)
  ),
  lonaxis = list(
    showgrid = TRUE,
    gridwidth = 0.5,
    range = c(-140, -55),
    dtick = 5
  ),
  lataxis = list(
    showgrid = TRUE,
    gridwidth = 0.5,
    range = c(20, 60),
    dtick = 5
  )
)

fig <- plot_geo(df, lat = ~Lat, lon = ~Lon, color = ~Globvalue)
fig <- fig %>% add_markers(
    text = ~paste(df$Globvalue, "inches"), hoverinfo = "text"
  )
fig <- fig %>% layout(title = 'US Precipitation 06-30-2015<br>Source: NOAA', geo = g)

fig
```
```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2015_06_30_precipitation.csv')

# change default color scale title
m <- list(colorbar = list(title = "Total Inches"))

# geo styling
g <- list(
  scope = 'north america',
  showland = TRUE,
  landcolor = toRGB("grey83"),
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white"),
  showlakes = TRUE,
  lakecolor = toRGB("white"),
  showsubunits = TRUE,
  showcountries = TRUE,
  resolution = 50,
  projection = list(
    type = 'conic conformal',
    rotation = list(lon = -100)
  ),
  lonaxis = list(
    showgrid = TRUE,
    gridwidth = 0.5,
    range = c(-140, -55),
    dtick = 5
  ),
  lataxis = list(
    showgrid = TRUE,
    gridwidth = 0.5,
    range = c(20, 60),
    dtick = 5
  )
)

fig <- plot_geo(df, lat = ~Lat, lon = ~Lon, color = ~Globvalue)
fig <- fig %>% add_markers(
    text = ~paste(df$Globvalue, "inches"), hoverinfo = "text"
  )
fig <- fig %>% layout(title = 'US Precipitation 06-30-2015<br>Source: NOAA', geo = g)

fig
```

## Mapbox Density
### Mapbox Access Token
```r
library(plotly)

quakes = read.csv('https://raw.githubusercontent.com/plotly/datasets/master/earthquakes-23k.csv')

fig <- quakes 
fig <- fig %>%
  plot_ly(
    type = 'densitymapbox',
    lat = ~Latitude,
    lon = ~Longitude,
    coloraxis = 'coloraxis',
    radius = 10) 
fig <- fig %>%
  layout(
    mapbox = list(
      style="stamen-terrain",
      center= list(lon=180)), coloraxis = list(colorscale = "Viridis"))

fig
```
```{r}
library(plotly)

quakes = read.csv('https://raw.githubusercontent.com/plotly/datasets/master/earthquakes-23k.csv')

fig <- quakes 
fig <- fig %>%
  plot_ly(
    type = 'densitymapbox',
    lat = ~Latitude,
    lon = ~Longitude,
    coloraxis = 'coloraxis',
    radius = 10) 
fig <- fig %>%
  layout(
    mapbox = list(
      style="stamen-terrain",
      center= list(lon=180)), coloraxis = list(colorscale = "Viridis"))

fig
```


## Mapbox Layers
### OpenStreetMap Tiles, no Token Needed
```r
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 

fig
```
```{r}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    marker = list(color = "fuchsia"),
    type = 'scattermapbox',
    hovertext = us_cities[,"City"]) 
fig <- fig %>%
  layout(
    mapbox = list(
      style = 'open-street-map',
      zoom =2.5,
      center = list(lon = -88, lat = 34))) 

fig
```
### Base Map Below A Trace: No Token Needed
```r
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    type = "scattermapbox",
    hovertext = us_cities[,"City"],
    marker = list(color = "fuchsia")) 
fig <- fig %>%
  layout(mapbox= list(
    style = "white-bg",
    zoom = 3,
    center = list(lon = -93 ,lat= 41),
    layers = list(list(
      below = 'traces',
      sourcetype = "raster",
      source = list("https://basemap.nationalmap.gov/arcgis/rest/services/USGSImageryOnly/MapServer/tile/{z}/{y}/{x}")))))

fig
```
```{r}
library(plotly)

us_cities = read.csv("https://raw.githubusercontent.com/plotly/datasets/master/us-cities-top-1k.csv")

fig <- us_cities 
fig <- fig %>%
  plot_ly(
    lat = ~lat,
    lon = ~lon,
    type = "scattermapbox",
    hovertext = us_cities[,"City"],
    marker = list(color = "fuchsia")) 
fig <- fig %>%
  layout(mapbox= list(
    style = "white-bg",
    zoom = 3,
    center = list(lon = -93 ,lat= 41),
    layers = list(list(
      below = 'traces',
      sourcetype = "raster",
      source = list("https://basemap.nationalmap.gov/arcgis/rest/services/USGSImageryOnly/MapServer/tile/{z}/{y}/{x}")))))

fig
```

## Bubble Maps
### United States Bubble Map
```r
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_us_cities.csv')

df$q <- with(df, cut(pop, quantile(pop)))
levels(df$q) <- paste(c("1st", "2nd", "3rd", "4th", "5th"), "Quantile")
df$q <- as.ordered(df$q)

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray85"),
  subunitwidth = 1,
  countrywidth = 1,
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white")
)

fig <- plot_geo(df, locationmode = 'USA-states', sizes = c(1, 250))
fig <- fig %>% add_markers(
    x = ~lon, y = ~lat, size = ~pop, color = ~q, hoverinfo = "text",
    text = ~paste(df$name, "<br />", df$pop/1e6, " million")
  )
fig <- fig %>% layout(title = '2014 US city populations<br>(Click legend to toggle)', geo = g)

fig
```
```{r}
library(plotly)
df <- read.csv('https://raw.githubusercontent.com/plotly/datasets/master/2014_us_cities.csv')

df$q <- with(df, cut(pop, quantile(pop)))
levels(df$q) <- paste(c("1st", "2nd", "3rd", "4th", "5th"), "Quantile")
df$q <- as.ordered(df$q)

g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray85"),
  subunitwidth = 1,
  countrywidth = 1,
  subunitcolor = toRGB("white"),
  countrycolor = toRGB("white")
)

fig <- plot_geo(df, locationmode = 'USA-states', sizes = c(1, 250))
fig <- fig %>% add_markers(
    x = ~lon, y = ~lat, size = ~pop, color = ~q, hoverinfo = "text",
    text = ~paste(df$name, "<br />", df$pop/1e6, " million")
  )
fig <- fig %>% layout(title = '2014 US city populations<br>(Click legend to toggle)', geo = g)

fig
```



<!--chapter:end:maps.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Artificial Intelligence and Machine Learning

<!--chapter:end:ml.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# 3D Charts
### Basic 3D Scatter Plot
```r
library(plotly)

mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

fig <- plot_ly(mtcars, x = ~wt, y = ~hp, z = ~qsec, color = ~am, colors = c('pink', 'blue'))
fig <- fig %>% add_markers(size = 10)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Weight'),
                     yaxis = list(title = 'Gross horsepower'),
                     zaxis = list(title = '1/4 mile time')))

fig
```

```{r}
library(plotly)

mtcars$am[which(mtcars$am == 0)] <- 'Automatic'
mtcars$am[which(mtcars$am == 1)] <- 'Manual'
mtcars$am <- as.factor(mtcars$am)

fig <- plot_ly(mtcars, x = ~wt, y = ~hp, z = ~qsec, color = ~am, colors = c('pink', 'blue'))
fig <- fig %>% add_markers(size = 10)
fig <- fig %>% layout(scene = list(xaxis = list(title = 'Weight'),
                     yaxis = list(title = 'Gross horsepower'),
                     zaxis = list(title = '1/4 mile time')))

fig
```

### Basic 3D Surface Plot
```r
# volcano is a numeric matrix that ships with R
fig <- plot_ly(z = ~volcano)
fig <- fig %>% add_surface()

fig
```

```{r}
# volcano is a numeric matrix that ships with R
fig <- plot_ly(z = ~volcano)
fig <- fig %>% add_surface()

fig
```


<!--chapter:end:3d.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Transforms
## Filter
### Basic Example
```r
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = mtcars$hp,
  y = mtcars$qsec,
  text = rownames(mtcars),
  hoverinfo = 'text',
  mode = 'markers',
  transforms = list(
    list(
      type = 'filter',
      target = 'y',
      operation = '>',
      value = mean(mtcars$qsec)
    )
  )
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = mtcars$hp,
  y = mtcars$qsec,
  text = rownames(mtcars),
  hoverinfo = 'text',
  mode = 'markers',
  transforms = list(
    list(
      type = 'filter',
      target = 'y',
      operation = '>',
      value = mean(mtcars$qsec)
    )
  )
)

fig
```


## Group
### Basic Example
```r
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = mtcars$hp,
  y = mtcars$qsec,
  text = paste("Make: ", rownames(mtcars),
               "<br>hp: ", mtcars$hp,
               "<br>qsec: ", mtcars$qsec,
               "<br>Cyl: ", mtcars$cyl),
  hoverinfo = 'text',
  mode = 'markers',
  transforms = list(
    list(
      type = 'groupby',
      groups = mtcars$cyl,
      styles = list(
        list(target = 4, value = list(marker =list(color = 'blue'))),
        list(target = 6, value = list(marker =list(color = 'red'))),
        list(target = 8, value = list(marker =list(color = 'black')))
      )
      )
    )
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = mtcars$hp,
  y = mtcars$qsec,
  text = paste("Make: ", rownames(mtcars),
               "<br>hp: ", mtcars$hp,
               "<br>qsec: ", mtcars$qsec,
               "<br>Cyl: ", mtcars$cyl),
  hoverinfo = 'text',
  mode = 'markers',
  transforms = list(
    list(
      type = 'groupby',
      groups = mtcars$cyl,
      styles = list(
        list(target = 4, value = list(marker =list(color = 'blue'))),
        list(target = 6, value = list(marker =list(color = 'red'))),
        list(target = 8, value = list(marker =list(color = 'black')))
      )
      )
    )
  )

fig
```


## Aggregation
### Basic Example
```r
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = diamonds$cut,
  y = diamonds$price,
  mode = 'markers',
  transforms = list(
    list(
      type = 'aggregate',
      groups = diamonds$cut,
      aggregations = list(
        list(
          target = 'y', func = 'sum', enabled = T
        )
      )
    )
  )
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatter',
  x = diamonds$cut,
  y = diamonds$price,
  mode = 'markers',
  transforms = list(
    list(
      type = 'aggregate',
      groups = diamonds$cut,
      aggregations = list(
        list(
          target = 'y', func = 'sum', enabled = T
        )
      )
    )
  )
)

fig
```
### Histogram Binning
```r
library(plotly)

df <- read.csv("https://plotly.com/~public.health/17.csv", skipNul = TRUE, encoding = "UTF-8")

labels <- function(size, label) {
  list(
    args = c("xbins.size", size), 
    label = label, 
    method = "restyle"
  )
}

fig <- df %>%
  plot_ly(
    x = ~date,
    autobinx = FALSE, 
    autobiny = TRUE, 
    marker = list(color = "rgb(68, 68, 68)"), 
    name = "date", 
    type = "histogram", 
    xbins = list(
      end = "2016-12-31 12:00", 
      size = "M1", 
      start = "1983-12-31 12:00"
    )
  )
fig <- fig %>% layout(
  paper_bgcolor = "rgb(240, 240, 240)", 
  plot_bgcolor = "rgb(240, 240, 240)", 
  title = "<b>Shooting Incidents</b><br>use dropdown to change bin size",
  xaxis = list(
    type = 'date'
  ),
  yaxis = list(
    title = "Incidents"
  ),
  updatemenus = list(
    list(
      x = 0.1, 
      y = 1.15,
      active = 1, 
      showactive = TRUE,
      buttons = list(
        labels("D1", "Day"),
        labels("M1", "Month"),
        labels("M6", "Half Year"),
        labels("M12", "Year")
      )
    )
  )
)

fig
```
```{r}
library(plotly)

df <- read.csv("https://plotly.com/~public.health/17.csv", skipNul = TRUE, encoding = "UTF-8")

labels <- function(size, label) {
  list(
    args = c("xbins.size", size), 
    label = label, 
    method = "restyle"
  )
}

fig <- df %>%
  plot_ly(
    x = ~date,
    autobinx = FALSE, 
    autobiny = TRUE, 
    marker = list(color = "rgb(68, 68, 68)"), 
    name = "date", 
    type = "histogram", 
    xbins = list(
      end = "2016-12-31 12:00", 
      size = "M1", 
      start = "1983-12-31 12:00"
    )
  )
fig <- fig %>% layout(
  paper_bgcolor = "rgb(240, 240, 240)", 
  plot_bgcolor = "rgb(240, 240, 240)", 
  title = "<b>Shooting Incidents</b><br>use dropdown to change bin size",
  xaxis = list(
    type = 'date'
  ),
  yaxis = list(
    title = "Incidents"
  ),
  updatemenus = list(
    list(
      x = 0.1, 
      y = 1.15,
      active = 1, 
      showactive = TRUE,
      buttons = list(
        labels("D1", "Day"),
        labels("M1", "Month"),
        labels("M6", "Half Year"),
        labels("M12", "Year")
      )
    )
  )
)

fig
```
### Mapping with Aggregations
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/worldhappiness.csv")

s <- schema()
agg <- s$transforms$aggregate$attributes$aggregations$items$aggregation$func$values


l = list()
for (i in 1:length(agg)) {
  ll = list(method = "restyle",
            args = list('transforms[0].aggregations[0].func', agg[i]),
            label = agg[i]) 
  l[[i]] = ll
}

fig <- df %>%
  plot_ly(
    type = 'choropleth',
    locationmode = 'country names',
    locations = ~Country,
    z = ~HappinessScore,
    autocolorscale = F,
    reversescale = T,
    colorscale = 'Portland', 
    transforms = list(list(
      type = 'aggregate',
      groups = ~Country,
      aggregations = list(
        list(target = 'z', func = 'sum', enabled = T)
      )
    ))
  )
fig <- fig %>% layout(
    title = "<b>World Happiness</b>",
    geo = list(
      showframe = F,
      showcoastlines = F
    ),
    updatemenus = list(
      list(
        x = 0.25,
        y = 1.04,
        xref = 'paper',
        yref = 'paper',
        yanchor = 'top',
        buttons = l
      )
    )
  )

fig
```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/worldhappiness.csv")

s <- schema()
agg <- s$transforms$aggregate$attributes$aggregations$items$aggregation$func$values


l = list()
for (i in 1:length(agg)) {
  ll = list(method = "restyle",
            args = list('transforms[0].aggregations[0].func', agg[i]),
            label = agg[i]) 
  l[[i]] = ll
}

fig <- df %>%
  plot_ly(
    type = 'choropleth',
    locationmode = 'country names',
    locations = ~Country,
    z = ~HappinessScore,
    autocolorscale = F,
    reversescale = T,
    colorscale = 'Portland', 
    transforms = list(list(
      type = 'aggregate',
      groups = ~Country,
      aggregations = list(
        list(target = 'z', func = 'sum', enabled = T)
      )
    ))
  )
fig <- fig %>% layout(
    title = "<b>World Happiness</b>",
    geo = list(
      showframe = F,
      showcoastlines = F
    ),
    updatemenus = list(
      list(
        x = 0.25,
        y = 1.04,
        xref = 'paper',
        yref = 'paper',
        yanchor = 'top',
        buttons = l
      )
    )
  )

fig
```


<!--chapter:end:transform.Rmd-->

```{r include=FALSE, cache=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE
)
```
# Animations

<!--chapter:end:animations.Rmd-->

