# Artificial Intelligence and Machine Learning
## ML Regression in R
### Linear Regression with R
```r
library(reshape2)
library(tidyverse)
library(tidymodels)
library(plotly)
data(tips)

y <- tips$tip
X <- tips$total_bill

set.seed(123)
tips_split <- initial_split(tips)
tips_training <- tips_split %>% 
  training()
tips_test <- tips_split %>% 
  testing()

lm_model <- linear_reg() %>% 
  set_engine('lm') %>% 
  set_mode('regression') %>%
  fit(tip ~ total_bill, data = tips_training) 

x_range <- seq(min(X), max(X), length.out = 100)
x_range <- matrix(x_range, nrow=100, ncol=1)
xdf <- data.frame(x_range)
colnames(xdf) <- c('total_bill')

ydf <- lm_model %>%
  predict(xdf) 

colnames(ydf) <- c('tip')
xy <- data.frame(xdf, ydf) 

fig <- plot_ly(data = tips_training, x = ~total_bill, y = ~tip, type = 'scatter', name = 'train', mode = 'markers', alpha = 0.65) %>% 
  add_trace(data = tips_test, x = ~total_bill, y = ~tip, type = 'scatter', name = 'test', mode = 'markers', alpha = 0.65 ) %>% 
  add_trace(data = xy, x = ~total_bill, y = ~tip, name = 'prediction', mode = 'lines', alpha = 1)
fig

```
```{r}
library(reshape2)
library(tidyverse)
library(tidymodels)
library(plotly)
data(tips)

y <- tips$tip
X <- tips$total_bill

set.seed(123)
tips_split <- initial_split(tips)
tips_training <- tips_split %>% 
  training()
tips_test <- tips_split %>% 
  testing()

lm_model <- linear_reg() %>% 
  set_engine('lm') %>% 
  set_mode('regression') %>%
  fit(tip ~ total_bill, data = tips_training) 

x_range <- seq(min(X), max(X), length.out = 100)
x_range <- matrix(x_range, nrow=100, ncol=1)
xdf <- data.frame(x_range)
colnames(xdf) <- c('total_bill')

ydf <- lm_model %>%
  predict(xdf) 

colnames(ydf) <- c('tip')
xy <- data.frame(xdf, ydf) 

fig <- plot_ly(data = tips_training, x = ~total_bill, y = ~tip, type = 'scatter', name = 'train', mode = 'markers', alpha = 0.65) %>% 
  add_trace(data = tips_test, x = ~total_bill, y = ~tip, type = 'scatter', name = 'test', mode = 'markers', alpha = 0.65 ) %>% 
  add_trace(data = xy, x = ~total_bill, y = ~tip, name = 'prediction', mode = 'lines', alpha = 1)
fig

```

## ROC and PR Curves
### ROC and PR Curves
```r
library(plotly)
library(tidymodels)
set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
data <- data.frame(X,y)
data$y <- as.factor(data$y)
X <- subset(data,select = -c(y))
logistic_glm <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  fit(y ~ ., data = data)

y_scores <- logistic_glm %>%
  predict(X, type = 'prob')

y_score <- y_scores$.pred_1
db <- data.frame(data$y, y_score)

z <- roc_curve(data = db, 'data.y', 'y_score')
z$specificity <- 1 - z$specificity
colnames(z) <- c('threshold', 'tpr', 'fpr')

fig1 <- plot_ly(x= y_score, color = data$y, colors = c('blue', 'red'), type = 'histogram', alpha = 0.5, nbinsx = 50) %>%
  layout(barmode = "overlay")
fig1
```
```{r}
library(plotly)
library(tidymodels)
set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
data <- data.frame(X,y)
data$y <- as.factor(data$y)
X <- subset(data,select = -c(y))
logistic_glm <-
  logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  fit(y ~ ., data = data)

y_scores <- logistic_glm %>%
  predict(X, type = 'prob')

y_score <- y_scores$.pred_1
db <- data.frame(data$y, y_score)

z <- roc_curve(data = db, 'data.y', 'y_score')
z$specificity <- 1 - z$specificity
colnames(z) <- c('threshold', 'tpr', 'fpr')

fig1 <- plot_ly(x= y_score, color = data$y, colors = c('blue', 'red'), type = 'histogram', alpha = 0.5, nbinsx = 50) %>%
  layout(barmode = "overlay")
fig1
```
```r
fig2 <- plot_ly(data = z, x = ~threshold) %>%
  add_trace(y = ~fpr, mode = 'lines', name = 'False Positive Rate', type = 'scatter')%>%
  add_trace(y = ~tpr, mode = 'lines', name = 'True Positive Rate', type = 'scatter')%>%
  layout(title = 'TPR and FPR at every threshold')
fig2 <- fig2 %>% layout(legend=list(title=list(text='<b> Rate </b>')))
fig2
```
```{r}
fig2 <- plot_ly(data = z, x = ~threshold) %>%
  add_trace(y = ~fpr, mode = 'lines', name = 'False Positive Rate', type = 'scatter')%>%
  add_trace(y = ~tpr, mode = 'lines', name = 'True Positive Rate', type = 'scatter')%>%
  layout(title = 'TPR and FPR at every threshold')
fig2 <- fig2 %>% layout(legend=list(title=list(text='<b> Rate </b>')))
fig2
```

### Basic binary ROC curve
```r
library(dplyr)
library(ggplot2)
library(plotly)
library(pROC)

set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
db <- data.frame(X,y)
db$y <- as.factor(db$y)
test_data = db[1:20]

model<- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  # Fit the model
  fit(y ~., data = db)


ypred <- predict(model,
                  new_data = test_data,
                  type = "prob")

yscore <- data.frame(ypred$.pred_0)
rdb <- cbind(db$y,yscore)
colnames(rdb) = c('y','yscore')


pdb <- roc_curve(rdb, y, yscore)
pdb$specificity <- 1 - pdb$specificity
auc = roc_auc(rdb, y, yscore)
auc = auc$.estimate

tit = paste('ROC Curve (AUC = ',toString(round(auc,2)),')',sep = '')


fig <-  plot_ly(data = pdb ,x =  ~specificity, y = ~sensitivity, type = 'scatter', mode = 'lines', fill = 'tozeroy') %>%
  layout(title = tit,xaxis = list(title = "False Positive Rate"), yaxis = list(title = "True Positive Rate")) %>%
add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'),inherit = FALSE, showlegend = FALSE)
fig

```

```{r}
library(dplyr)
library(ggplot2)
library(plotly)
library(pROC)

set.seed(0)
X <- matrix(rnorm(10000),nrow=500)
y <- sample(0:1, 500, replace=TRUE)
db <- data.frame(X,y)
db$y <- as.factor(db$y)
test_data = db[1:20]

model<- logistic_reg() %>%
  set_engine("glm") %>%
  set_mode("classification") %>%
  # Fit the model
  fit(y ~., data = db)


ypred <- predict(model,
                  new_data = test_data,
                  type = "prob")

yscore <- data.frame(ypred$.pred_0)
rdb <- cbind(db$y,yscore)
colnames(rdb) = c('y','yscore')


pdb <- roc_curve(rdb, y, yscore)
pdb$specificity <- 1 - pdb$specificity
auc = roc_auc(rdb, y, yscore)
auc = auc$.estimate

tit = paste('ROC Curve (AUC = ',toString(round(auc,2)),')',sep = '')


fig <-  plot_ly(data = pdb ,x =  ~specificity, y = ~sensitivity, type = 'scatter', mode = 'lines', fill = 'tozeroy') %>%
  layout(title = tit,xaxis = list(title = "False Positive Rate"), yaxis = list(title = "True Positive Rate")) %>%
add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'),inherit = FALSE, showlegend = FALSE)
fig

```
### Multiclass ROC Curve
```r
library(plotly)
library(tidymodels)
library(fastDummies)

# Artificially add noise to make task harder
data(iris)
ind <- sample.int(150, 50)
samples <- sample(x = iris$Species, size = 50)
iris[ind,'Species'] = samples

# Define the inputs and outputs
X <- subset(iris, select = -c(Species))
iris$Species <- as.factor(iris$Species)

# Fit the model
logistic <-
  multinom_reg() %>%
  set_engine("nnet") %>%
  set_mode("classification") %>%
  fit(Species ~ ., data = iris)

y_scores <- logistic %>%
  predict(X, type = 'prob')

# One hot encode the labels in order to plot them
y_onehot <- dummy_cols(iris$Species)
colnames(y_onehot) <- c('drop', 'setosa', 'versicolor', 'virginica')
y_onehot <- subset(y_onehot, select = -c(drop))

z = cbind(y_scores, y_onehot)

z$setosa <- as.factor(z$setosa)
roc_setosa <- roc_curve(data = z, setosa, .pred_setosa)
roc_setosa$specificity <- 1 - roc_setosa$specificity
colnames(roc_setosa) <- c('threshold', 'tpr', 'fpr')
auc_setosa <- roc_auc(data = z, setosa, .pred_setosa)
auc_setosa <- auc_setosa$.estimate
setosa <- paste('setosa (AUC=',toString(round(1-auc_setosa,2)),')',sep = '')

z$versicolor <- as.factor(z$versicolor)
roc_versicolor <- roc_curve(data = z, versicolor, .pred_versicolor)
roc_versicolor$specificity <- 1 - roc_versicolor$specificity
colnames(roc_versicolor) <- c('threshold', 'tpr', 'fpr')
auc_versicolor <- roc_auc(data = z, versicolor, .pred_versicolor)
auc_versicolor <- auc_versicolor$.estimate
versicolor <- paste('versicolor (AUC=',toString(round(1-auc_versicolor,2)),')', sep = '')

z$virginica <- as.factor(z$virginica)
roc_virginica <- roc_curve(data = z, virginica, .pred_virginica)
roc_virginica$specificity <- 1 - roc_virginica$specificity
colnames(roc_virginica) <- c('threshold', 'tpr', 'fpr')
auc_virginica <- roc_auc(data = z, virginica, .pred_virginica)
auc_virginica <- auc_virginica$.estimate
virginica <- paste('virginica (AUC=',toString(round(1-auc_virginica,2)),')',sep = '')

# Create an empty figure, and iteratively add a line for each class
fig <- plot_ly()%>%
  add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'), showlegend = FALSE) %>%
  add_trace(data = roc_setosa,x = ~fpr, y = ~tpr, mode = 'lines', name = setosa, type = 'scatter')%>%
  add_trace(data = roc_versicolor,x = ~fpr, y = ~tpr, mode = 'lines', name = versicolor, type = 'scatter')%>%
  add_trace(data = roc_virginica,x = ~fpr, y = ~tpr, mode = 'lines', name = virginica, type = 'scatter')%>%
  layout(xaxis = list(
    title = "False Positive Rate"
  ), yaxis = list(
    title = "True Positive Rate"
  ),legend = list(x = 100, y = 0.5))
fig

```
```{r}
library(plotly)
library(tidymodels)
library(fastDummies)

# Artificially add noise to make task harder
data(iris)
ind <- sample.int(150, 50)
samples <- sample(x = iris$Species, size = 50)
iris[ind,'Species'] = samples

# Define the inputs and outputs
X <- subset(iris, select = -c(Species))
iris$Species <- as.factor(iris$Species)

# Fit the model
logistic <-
  multinom_reg() %>%
  set_engine("nnet") %>%
  set_mode("classification") %>%
  fit(Species ~ ., data = iris)

y_scores <- logistic %>%
  predict(X, type = 'prob')

# One hot encode the labels in order to plot them
y_onehot <- dummy_cols(iris$Species)
colnames(y_onehot) <- c('drop', 'setosa', 'versicolor', 'virginica')
y_onehot <- subset(y_onehot, select = -c(drop))

z = cbind(y_scores, y_onehot)

z$setosa <- as.factor(z$setosa)
roc_setosa <- roc_curve(data = z, setosa, .pred_setosa)
roc_setosa$specificity <- 1 - roc_setosa$specificity
colnames(roc_setosa) <- c('threshold', 'tpr', 'fpr')
auc_setosa <- roc_auc(data = z, setosa, .pred_setosa)
auc_setosa <- auc_setosa$.estimate
setosa <- paste('setosa (AUC=',toString(round(1-auc_setosa,2)),')',sep = '')

z$versicolor <- as.factor(z$versicolor)
roc_versicolor <- roc_curve(data = z, versicolor, .pred_versicolor)
roc_versicolor$specificity <- 1 - roc_versicolor$specificity
colnames(roc_versicolor) <- c('threshold', 'tpr', 'fpr')
auc_versicolor <- roc_auc(data = z, versicolor, .pred_versicolor)
auc_versicolor <- auc_versicolor$.estimate
versicolor <- paste('versicolor (AUC=',toString(round(1-auc_versicolor,2)),')', sep = '')

z$virginica <- as.factor(z$virginica)
roc_virginica <- roc_curve(data = z, virginica, .pred_virginica)
roc_virginica$specificity <- 1 - roc_virginica$specificity
colnames(roc_virginica) <- c('threshold', 'tpr', 'fpr')
auc_virginica <- roc_auc(data = z, virginica, .pred_virginica)
auc_virginica <- auc_virginica$.estimate
virginica <- paste('virginica (AUC=',toString(round(1-auc_virginica,2)),')',sep = '')

# Create an empty figure, and iteratively add a line for each class
fig <- plot_ly()%>%
  add_segments(x = 0, xend = 1, y = 0, yend = 1, line = list(dash = "dash", color = 'black'), showlegend = FALSE) %>%
  add_trace(data = roc_setosa,x = ~fpr, y = ~tpr, mode = 'lines', name = setosa, type = 'scatter')%>%
  add_trace(data = roc_versicolor,x = ~fpr, y = ~tpr, mode = 'lines', name = versicolor, type = 'scatter')%>%
  add_trace(data = roc_virginica,x = ~fpr, y = ~tpr, mode = 'lines', name = virginica, type = 'scatter')%>%
  layout(xaxis = list(
    title = "False Positive Rate"
  ), yaxis = list(
    title = "True Positive Rate"
  ),legend = list(x = 100, y = 0.5))
fig

```

## PCA Visulization
### Visualize all the original dimensions
```r
library(plotly)

data(iris)

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4,
            titlefont=list(size=13))


fig <- iris %>%
  plot_ly()
fig <- fig %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label='sepal length', values=~Sepal.Length),
      list(label='sepal width', values=~Sepal.Width),
      list(label='petal length', values=~Petal.Length),
      list(label='petal width', values=~Petal.Width)
    ),
    color = ~Species, colors = c('#636EFA','#EF553B','#00CC96') ,
    marker = list(
      size = 7,
      line = list(
        width = 1,
        color = 'rgb(230,230,230)'
      )
    )
  )
fig <-  fig %>% style(diagonal = list(visible = FALSE))
fig <- fig %>%
  layout(
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )

fig

```
```{r}
library(plotly)

data(iris)

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4,
            titlefont=list(size=13))


fig <- iris %>%
  plot_ly()
fig <- fig %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label='sepal length', values=~Sepal.Length),
      list(label='sepal width', values=~Sepal.Width),
      list(label='petal length', values=~Petal.Length),
      list(label='petal width', values=~Petal.Width)
    ),
    color = ~Species, colors = c('#636EFA','#EF553B','#00CC96') ,
    marker = list(
      size = 7,
      line = list(
        width = 1,
        color = 'rgb(230,230,230)'
      )
    )
  )
fig <-  fig %>% style(diagonal = list(visible = FALSE))
fig <- fig %>%
  layout(
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )

fig

```

### Visualize all the principal components
```r
library(plotly)
library(stats)
data(iris)
X <- subset(iris, select = -c(Species))
prin_comp <- prcomp(X)
explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
explained_variance_ratio <- 100 * explained_variance_ratio
components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, iris$Species)
components$PC3 <- -components$PC3
components$PC2 <- -components$PC2

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4,
            titlefont=list(size=13))

fig <- components %>%
  plot_ly()  %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label=paste('PC 1 (',toString(round(explained_variance_ratio[1],1)),'%)',sep = ''), values=~PC1),
      list(label=paste('PC 2 (',toString(round(explained_variance_ratio[2],1)),'%)',sep = ''), values=~PC2),
      list(label=paste('PC 3 (',toString(round(explained_variance_ratio[3],1)),'%)',sep = ''), values=~PC3),
      list(label=paste('PC 4 (',toString(round(explained_variance_ratio[4],1)),'%)',sep = ''), values=~PC4)
    ),
    color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96')
  ) %>%
  style(diagonal = list(visible = FALSE)) %>%
  layout(
    legend=list(title=list(text='color')),
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )

fig

```
```{r}
library(plotly)
library(stats)
data(iris)
X <- subset(iris, select = -c(Species))
prin_comp <- prcomp(X)
explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
explained_variance_ratio <- 100 * explained_variance_ratio
components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, iris$Species)
components$PC3 <- -components$PC3
components$PC2 <- -components$PC2

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4,
            titlefont=list(size=13))

fig <- components %>%
  plot_ly()  %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label=paste('PC 1 (',toString(round(explained_variance_ratio[1],1)),'%)',sep = ''), values=~PC1),
      list(label=paste('PC 2 (',toString(round(explained_variance_ratio[2],1)),'%)',sep = ''), values=~PC2),
      list(label=paste('PC 3 (',toString(round(explained_variance_ratio[3],1)),'%)',sep = ''), values=~PC3),
      list(label=paste('PC 4 (',toString(round(explained_variance_ratio[4],1)),'%)',sep = ''), values=~PC4)
    ),
    color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96')
  ) %>%
  style(diagonal = list(visible = FALSE)) %>%
  layout(
    legend=list(title=list(text='color')),
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )

fig

```
### Visualize a subset of the principal components
```r
library(plotly)
library(stats)
library(MASS)

db = Boston

prin_comp <- prcomp(db, rank. = 4)

components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, db$medv)
components$PC2 <- -components$PC2
colnames(components)[5] = 'Median_Price'

tot_explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
tot_explained_variance_ratio <- 100 * sum(tot_explained_variance_ratio)

tit = 'Total Explained Variance = 99.56'

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4)

fig <- components %>%
  plot_ly() %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label='PC1', values=~PC1),
      list(label='PC2', values=~PC2),
      list(label='PC3', values=~PC3),
      list(label='PC4', values=~PC4)
    ),
    color=~Median_Price,
    marker = list(
      size = 7
    )
  ) %>% style(diagonal = list(visible = F)) %>%
  layout(
    title= tit,
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )
options(warn=-1)
fig

```
```{r}
library(plotly)
library(stats)
library(MASS)

db = Boston

prin_comp <- prcomp(db, rank. = 4)

components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, db$medv)
components$PC2 <- -components$PC2
colnames(components)[5] = 'Median_Price'

tot_explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
tot_explained_variance_ratio <- 100 * sum(tot_explained_variance_ratio)

tit = 'Total Explained Variance = 99.56'

axis = list(showline=FALSE,
            zeroline=FALSE,
            gridcolor='#ffff',
            ticklen=4)

fig <- components %>%
  plot_ly() %>%
  add_trace(
    type = 'splom',
    dimensions = list(
      list(label='PC1', values=~PC1),
      list(label='PC2', values=~PC2),
      list(label='PC3', values=~PC3),
      list(label='PC4', values=~PC4)
    ),
    color=~Median_Price,
    marker = list(
      size = 7
    )
  ) %>% style(diagonal = list(visible = F)) %>%
  layout(
    title= tit,
    hovermode='closest',
    dragmode= 'select',
    plot_bgcolor='rgba(240,240,240, 0.95)',
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4),
    xaxis2=axis,
    xaxis3=axis,
    xaxis4=axis,
    yaxis2=axis,
    yaxis3=axis,
    yaxis4=axis
  )
options(warn=-1)
fig

```
### Visualize PCA with scatter3d
```r
data("iris")

X <- subset(iris, select = -c(Species))

prin_comp <- prcomp(X, rank. = 3)

components <- prin_comp[["x"]]
components <- data.frame(components)
components$PC2 <- -components$PC2
components$PC3 <- -components$PC3
components = cbind(components, iris$Species)

tot_explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
tot_explained_variance_ratio <- 100 * sum(tot_explained_variance_ratio)

tit = 'Total Explained Variance = 99.48'

fig <- plot_ly(components, x = ~PC1, y = ~PC2, z = ~PC3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96') ) %>%
  add_markers(size = 12)


fig <- fig %>%
  layout(
    title = tit,
    scene = list(bgcolor = "#e5ecf6")
)

fig

```
```{r}
data("iris")

X <- subset(iris, select = -c(Species))

prin_comp <- prcomp(X, rank. = 3)

components <- prin_comp[["x"]]
components <- data.frame(components)
components$PC2 <- -components$PC2
components$PC3 <- -components$PC3
components = cbind(components, iris$Species)

tot_explained_variance_ratio <- summary(prin_comp)[["importance"]]['Proportion of Variance',]
tot_explained_variance_ratio <- 100 * sum(tot_explained_variance_ratio)

tit = 'Total Explained Variance = 99.48'

fig <- plot_ly(components, x = ~PC1, y = ~PC2, z = ~PC3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96') ) %>%
  add_markers(size = 12)


fig <- fig %>%
  layout(
    title = tit,
    scene = list(bgcolor = "#e5ecf6")
)

fig

```
### Visualize Loadings
```r
library(plotly)
library(stats)
data(iris)
X <- subset(iris, select = -c(Species))
prin_comp <- prcomp(X, rank = 2)
components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, iris$Species)
components$PC2 <- -components$PC2
explained_variance <- summary(prin_comp)[["sdev"]]
explained_variance <- explained_variance[1:2]
comp <- prin_comp[["rotation"]]
comp[,'PC2'] <- - comp[,'PC2']
loadings <- comp
for (i in seq(explained_variance)){
  loadings[,i] <- comp[,i] * explained_variance[i]
}

features = c('sepal_length', 'sepal_width', 'petal_length', 'petal_width')

fig <- plot_ly(components, x = ~PC1, y = ~PC2, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96'), type = 'scatter', mode = 'markers') %>%
  layout(
    legend=list(title=list(text='color')),
    plot_bgcolor = "#e5ecf6",
    xaxis = list(
      title = "0"),
    yaxis = list(
      title = "1"))
for (i in seq(4)){
  fig <- fig %>%
    add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'black'),inherit = FALSE, showlegend = FALSE) %>%
    add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = features[i], xanchor = 'center', yanchor= 'bottom')
}

fig

```

```{r}
library(plotly)
library(stats)
data(iris)
X <- subset(iris, select = -c(Species))
prin_comp <- prcomp(X, rank = 2)
components <- prin_comp[["x"]]
components <- data.frame(components)
components <- cbind(components, iris$Species)
components$PC2 <- -components$PC2
explained_variance <- summary(prin_comp)[["sdev"]]
explained_variance <- explained_variance[1:2]
comp <- prin_comp[["rotation"]]
comp[,'PC2'] <- - comp[,'PC2']
loadings <- comp
for (i in seq(explained_variance)){
  loadings[,i] <- comp[,i] * explained_variance[i]
}

features = c('sepal_length', 'sepal_width', 'petal_length', 'petal_width')

fig <- plot_ly(components, x = ~PC1, y = ~PC2, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96'), type = 'scatter', mode = 'markers') %>%
  layout(
    legend=list(title=list(text='color')),
    plot_bgcolor = "#e5ecf6",
    xaxis = list(
      title = "0"),
    yaxis = list(
      title = "1"))
for (i in seq(4)){
  fig <- fig %>%
    add_segments(x = 0, xend = loadings[i, 1], y = 0, yend = loadings[i, 2], line = list(color = 'black'),inherit = FALSE, showlegend = FALSE) %>%
    add_annotations(x=loadings[i, 1], y=loadings[i, 2], ax = 0, ay = 0,text = features[i], xanchor = 'center', yanchor= 'bottom')
}

fig

```
## t-SNE and UMAP projections
### Basic t-SNE projections
```r
library(plotly) 
library(stats) 
data(iris) 
X <- subset(iris, select = -c(Species)) 
axis = list(showline=FALSE, 
            zeroline=FALSE, 
            gridcolor='#ffff', 
            ticklen=4)
fig <- iris %>%  
  plot_ly()  %>%  
  add_trace(  
    type = 'splom',  
    dimensions = list( 
      list(label = 'sepal_width',values=~Sepal.Width),  
      list(label = 'sepal_length',values=~Sepal.Length),  
      list(label ='petal_width',values=~Petal.Width),  
      list(label = 'petal_length',values=~Petal.Length)),  
    color = ~Species, colors = c('#636EFA','#EF553B','#00CC96') 
  ) 
fig <- fig %>% 
  layout( 
    legend=list(title=list(text='species')), 
    hovermode='closest', 
    dragmode= 'select', 
    plot_bgcolor='rgba(240,240,240,0.95)', 
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    xaxis2=axis, 
    xaxis3=axis, 
    xaxis4=axis, 
    yaxis2=axis, 
    yaxis3=axis, 
    yaxis4=axis 
  ) 
fig

```
```{r}
library(plotly) 
library(stats) 
data(iris) 
X <- subset(iris, select = -c(Species)) 
axis = list(showline=FALSE, 
            zeroline=FALSE, 
            gridcolor='#ffff', 
            ticklen=4)
fig <- iris %>%  
  plot_ly()  %>%  
  add_trace(  
    type = 'splom',  
    dimensions = list( 
      list(label = 'sepal_width',values=~Sepal.Width),  
      list(label = 'sepal_length',values=~Sepal.Length),  
      list(label ='petal_width',values=~Petal.Width),  
      list(label = 'petal_length',values=~Petal.Length)),  
    color = ~Species, colors = c('#636EFA','#EF553B','#00CC96') 
  ) 
fig <- fig %>% 
  layout( 
    legend=list(title=list(text='species')), 
    hovermode='closest', 
    dragmode= 'select', 
    plot_bgcolor='rgba(240,240,240,0.95)', 
    xaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    yaxis=list(domain=NULL, showline=F, zeroline=F, gridcolor='#ffff', ticklen=4), 
    xaxis2=axis, 
    xaxis3=axis, 
    xaxis4=axis, 
    yaxis2=axis, 
    yaxis3=axis, 
    yaxis4=axis 
  ) 
fig

```
### Project data into 2D with t-SNE
```r
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

set.seed(0)
tsne <- tsne(features, initial_dims = 2)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, type = 'scatter', mode = 'markers', split = ~iris$Species)

fig <- fig %>%
  layout(
    plot_bgcolor = "#e5ecf6"
  )

fig

```
```{r}
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

set.seed(0)
tsne <- tsne(features, initial_dims = 2)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, type = 'scatter', mode = 'markers', split = ~iris$Species)

fig <- fig %>%
  layout(
    plot_bgcolor = "#e5ecf6"
  )

fig


```
### Project data into 3D with t-SNE
```r
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

#set.seed(0)
tsne <- tsne(features, initial_dims = 3, k =3)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96') ) %>% 
  add_markers(size = 8) %>%
  layout( 
    xaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'), 
    yaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'),
    scene =list(bgcolor = "#e5ecf6"))
fig

```
```{r}
library(tsne)
library(plotly)
data("iris")

features <- subset(iris, select = -c(Species)) 

#set.seed(0)
tsne <- tsne(features, initial_dims = 3, k =3)
tsne <- data.frame(tsne)
pdb <- cbind(tsne,iris$Species)
options(warn = -1)
fig <-  plot_ly(data = pdb ,x =  ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96') ) %>% 
  add_markers(size = 8) %>%
  layout( 
    xaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'), 
    yaxis = list(
      zerolinecolor = "#ffff",
      zerolinewidth = 2,
      gridcolor='#ffff'),
    scene =list(bgcolor = "#e5ecf6"))
fig

```
### Projections with UMAP
```r
library(plotly) 
library(umap) 
iris.data = iris[, grep("Sepal|Petal", colnames(iris))] 
iris.labels = iris[, "Species"] 
iris.umap = umap(iris.data, n_components = 2, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 

fig <- plot_ly(final, x = ~X1, y = ~X2, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96'), type = 'scatter', mode = 'markers')%>%  
  layout(
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='species')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 

iris.umap = umap(iris.data, n_components = 3, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 

fig2 <- plot_ly(final, x = ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96')) 
fig2 <- fig2 %>% add_markers() 
fig2 <- fig2 %>% layout(scene = list(xaxis = list(title = '0'), 
                                     yaxis = list(title = '1'), 
                                     zaxis = list(title = '2'))) 

fig 

```
```{r}
library(plotly) 
library(umap) 
iris.data = iris[, grep("Sepal|Petal", colnames(iris))] 
iris.labels = iris[, "Species"] 
iris.umap = umap(iris.data, n_components = 2, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 

fig <- plot_ly(final, x = ~X1, y = ~X2, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96'), type = 'scatter', mode = 'markers')%>%  
  layout(
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='species')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 

iris.umap = umap(iris.data, n_components = 3, random_state = 15) 
layout <- iris.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, iris$Species) 

fig2 <- plot_ly(final, x = ~X1, y = ~X2, z = ~X3, color = ~iris$Species, colors = c('#636EFA','#EF553B','#00CC96')) 
fig2 <- fig2 %>% add_markers() 
fig2 <- fig2 %>% layout(scene = list(xaxis = list(title = '0'), 
                                     yaxis = list(title = '1'), 
                                     zaxis = list(title = '2'))) 

fig 

```
```r
fig2
```
```{r}
fig2
```
### Visualizing image datasets
```r
library(rsvd) 
library(plotly) 
library(umap) 
data('digits') 
digits.data = digits[, grep("pixel", colnames(digits))] 
digits.labels = digits[, "label"] 
digits.umap = umap(digits.data, n_components = 2, k = 10) 
layout <- digits.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, digits[,'label']) 
colnames(final) <- c('X1', 'X2', 'label') 

fig <- plot_ly(final, x = ~X1, y = ~X2, split = ~label,  type = 'scatter', mode = 'markers')%>%  
  layout(  
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='digit')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 
fig

```
```{r}
library(rsvd) 
library(plotly) 
library(umap) 
data('digits') 
digits.data = digits[, grep("pixel", colnames(digits))] 
digits.labels = digits[, "label"] 
digits.umap = umap(digits.data, n_components = 2, k = 10) 
layout <- digits.umap[["layout"]] 
layout <- data.frame(layout) 
final <- cbind(layout, digits[,'label']) 
colnames(final) <- c('X1', 'X2', 'label') 

fig <- plot_ly(final, x = ~X1, y = ~X2, split = ~label,  type = 'scatter', mode = 'markers')%>%  
  layout(  
    plot_bgcolor = "#e5ecf6",
    legend=list(title=list(text='digit')), 
    xaxis = list( 
      title = "0"),  
    yaxis = list( 
      title = "1")) 
fig

```

