# Scientific Charts
## Log Plots
### Log Axes
```r
library(plotly)
d <- diamonds[sample(nrow(diamonds), 1000), ]
# without log scales
fig <- plot_ly(d, x = ~carat, y = ~price) %>% add_markers()

fig
```
```{r}
library(plotly)
d <- diamonds[sample(nrow(diamonds), 1000), ]
# without log scales
fig <- plot_ly(d, x = ~carat, y = ~price) %>% add_markers()

fig
```
```r
# with log scales
fig <- layout(fig, xaxis = list(type = "log"),
       yaxis = list(type = "log"))

fig
```
```{r}
# with log scales
fig <- layout(fig, xaxis = list(type = "log"),
       yaxis = list(type = "log"))

fig
```

## Contour Plots
### Basic Contour
```r
library(plotly)

fig <- plot_ly(z = ~volcano, type = "contour")

fig
```
```{r}
library(plotly)

fig <- plot_ly(z = ~volcano, type = "contour")

fig
```
### Smoothing Contour Coloring
```r
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z = matrix(c(10, 10.625, 12.5, 15.625, 20, 5.625, 6.25, 8.125, 
               11.25, 15.625, 2.5, 3.125, 5, 8.125, 12.5, 0.625, 
               1.25, 3.125, 6.25, 10.625, 0, 0.625, 2.5, 5.625, 
               10), nrow=5, ncol=5),
  contours = list(
    coloring = 'heatmap'
  )
)

fig

```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'contour',
  z = matrix(c(10, 10.625, 12.5, 15.625, 20, 5.625, 6.25, 8.125, 
               11.25, 15.625, 2.5, 3.125, 5, 8.125, 12.5, 0.625, 
               1.25, 3.125, 6.25, 10.625, 0, 0.625, 2.5, 5.625, 
               10), nrow=5, ncol=5),
  contours = list(
    coloring = 'heatmap'
  )
)

fig

```

## Heatmaps
### Basic Heatmap
```r
library(plotly)
fig <- plot_ly(z = volcano, type = "heatmap")

fig
```
```{r}
library(plotly)
fig <- plot_ly(z = volcano, type = "heatmap")

fig
```
### Sequential Colorscales: Greys
```r
fig <- plot_ly(z = volcano, colors = "Greys", type = "heatmap")

fig
```
```{r}
fig <- plot_ly(z = volcano, colors = "Greys", type = "heatmap")

fig
```

## Parallel Coordinates
### Basic Parallel Cordinates Plots
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/iris.csv")

fig <- df %>% plot_ly(type = 'parcoords',
          line = list(color = ~species_id,
                      colorscale = list(c(0,'red'),c(0.5,'green'),c(1,'blue'))),
          dimensions = list(
            list(range = c(2,4.5),
                 label = 'Sepal Width', values = ~sepal_width),
            list(range = c(4,8),
                 constraintrange = c(5,6),
                 label = 'Sepal Length', values = ~sepal_length),
            list(range = c(0,2.5),
                 label = 'Petal Width', values = ~petal_width),
            list(range = c(1,7),
                 label = 'Petal Length', values = ~petal_length)
            )
          )

fig
```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/iris.csv")

fig <- df %>% plot_ly(type = 'parcoords',
          line = list(color = ~species_id,
                      colorscale = list(c(0,'red'),c(0.5,'green'),c(1,'blue'))),
          dimensions = list(
            list(range = c(2,4.5),
                 label = 'Sepal Width', values = ~sepal_width),
            list(range = c(4,8),
                 constraintrange = c(5,6),
                 label = 'Sepal Length', values = ~sepal_length),
            list(range = c(0,2.5),
                 label = 'Petal Width', values = ~petal_width),
            list(range = c(1,7),
                 label = 'Petal Length', values = ~petal_length)
            )
          )

fig
```
### Advanced Parallel Coordinates Plot
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/parcoords_data.csv")

fig <- df %>%
  plot_ly(width = 1000, height = 600) 
fig <- fig %>% add_trace(type = 'parcoords',
          line = list(color = ~colorVal,
                      colorscale = 'Jet',
                      showscale = TRUE,
                      reversescale = TRUE,
                      cmin = -4000,
                      cmax = -100),
          dimensions = list(
            list(range = c(~min(blockHeight),~max(blockHeight)),
                 constraintrange = c(100000,150000),
                 label = 'Block Height', values = ~blockHeight),
            list(range = c(~min(blockWidth),~max(blockWidth)),
                 label = 'Block Width', values = ~blockWidth),
            list(tickvals = c(0,0.5,1,2,3),
                 ticktext = c('A','AB','B','Y','Z'),
                 label = 'Cyclinder Material', values = ~cycMaterial),
            list(range = c(-1,4),
                 tickvals = c(0,1,2,3),
                 label = 'Block Material', values = ~blockMaterial),
            list(range = c(~min(totalWeight),~max(totalWeight)),
                 visible = TRUE,
                 label = 'Total Weight', values = ~totalWeight),
            list(range = c(~min(assemblyPW),~max(assemblyPW)),
                 label = 'Assembly Penalty Weight', values = ~assemblyPW),
            list(range = c(~min(HstW),~max(HstW)),
                 label = 'Height st Width', values = ~HstW),
            list(range = c(~min(minHW),~max(minHW)),
                 label = 'Min Height Width', values = ~minHW),
            list(range = c(~min(minWD),~max(minWD)),
                 label = 'Min Width Diameter', values = ~minWD),
            list(range = c(~min(rfBlock),~max(rfBlock)),
                 label = 'RF Block', values = ~rfBlock)
            )
          )


fig

```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/bcdunbar/datasets/master/parcoords_data.csv")

fig <- df %>%
  plot_ly(width = 1000, height = 600) 
fig <- fig %>% add_trace(type = 'parcoords',
          line = list(color = ~colorVal,
                      colorscale = 'Jet',
                      showscale = TRUE,
                      reversescale = TRUE,
                      cmin = -4000,
                      cmax = -100),
          dimensions = list(
            list(range = c(~min(blockHeight),~max(blockHeight)),
                 constraintrange = c(100000,150000),
                 label = 'Block Height', values = ~blockHeight),
            list(range = c(~min(blockWidth),~max(blockWidth)),
                 label = 'Block Width', values = ~blockWidth),
            list(tickvals = c(0,0.5,1,2,3),
                 ticktext = c('A','AB','B','Y','Z'),
                 label = 'Cyclinder Material', values = ~cycMaterial),
            list(range = c(-1,4),
                 tickvals = c(0,1,2,3),
                 label = 'Block Material', values = ~blockMaterial),
            list(range = c(~min(totalWeight),~max(totalWeight)),
                 visible = TRUE,
                 label = 'Total Weight', values = ~totalWeight),
            list(range = c(~min(assemblyPW),~max(assemblyPW)),
                 label = 'Assembly Penalty Weight', values = ~assemblyPW),
            list(range = c(~min(HstW),~max(HstW)),
                 label = 'Height st Width', values = ~HstW),
            list(range = c(~min(minHW),~max(minHW)),
                 label = 'Min Height Width', values = ~minHW),
            list(range = c(~min(minWD),~max(minWD)),
                 label = 'Min Width Diameter', values = ~minWD),
            list(range = c(~min(rfBlock),~max(rfBlock)),
                 label = 'RF Block', values = ~rfBlock)
            )
          )


fig

```
## Polar Charts
### Basic Polar Charts
```r
library(plotly)

fig <- plot_ly(
  type = 'scatterpolar',
  r = c(0,1,2,2),
  theta = c(0,45,90,0),
  mode = 'markers'
)

fig
```
```{r}
library(plotly)

fig <- plot_ly(
  type = 'scatterpolar',
  r = c(0,1,2,2),
  theta = c(0,45,90,0),
  mode = 'markers'
)

fig
```
### Line Polar Charts
```r
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/polar_dataset.csv")

fig <- plot_ly(
  df,
  type = 'scatterpolar',
  mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = ~x1,
    theta = ~y,
    name = 'Figure8',
    line = list(
      color = 'peru'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x2,
    theta = ~y,
    name = 'Cardioid',
    line = list(
      color = 'darkviolet'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x3,
    theta = ~y,
    name = 'Hypercardioid',
    line = list(
      color = 'deepskyblue'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x4,
    theta = ~y,
    name = 'Subcardioid',
    line = list(
      color = 'orangered'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x5,
    theta = ~y,
    name = 'Supercardioid',
    line = list(
      color = 'green'
    )
  ) 

fig <- fig %>%
  layout(
    title = 'Mic Patterns',
    font = list(
      family = 'Arial',
      size = 12,
      color = '#000'
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

df <- read.csv("https://raw.githubusercontent.com/plotly/datasets/master/polar_dataset.csv")

fig <- plot_ly(
  df,
  type = 'scatterpolar',
  mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = ~x1,
    theta = ~y,
    name = 'Figure8',
    line = list(
      color = 'peru'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x2,
    theta = ~y,
    name = 'Cardioid',
    line = list(
      color = 'darkviolet'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x3,
    theta = ~y,
    name = 'Hypercardioid',
    line = list(
      color = 'deepskyblue'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x4,
    theta = ~y,
    name = 'Subcardioid',
    line = list(
      color = 'orangered'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = ~x5,
    theta = ~y,
    name = 'Supercardioid',
    line = list(
      color = 'green'
    )
  ) 

fig <- fig %>%
  layout(
    title = 'Mic Patterns',
    font = list(
      family = 'Arial',
      size = 12,
      color = '#000'
    ),
    showlegend = F
  )

fig
```
### Area Polar Area
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 1.5, 1.5, 0, 2.5, 2.5, 0),
    theta = c(0, 10, 25, 0, 205, 215, 0),
    fill = 'toself',
    fillcolor = '#709Bff',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3.5, 3.5, 0),
    theta = c(0, 55, 75, 0),
    fill = 'toself',
    fillcolor = '#E4FF87',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4.5, 4.5, 0, 4.5, 4.5, 0),
    theta = c(0, 100, 120, 0, 305, 320, 0),
    fill = 'toself',
    fillcolor = '#FFAA70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4, 4, 0),
    theta = c(0, 165, 195, 0),
    fill = 'toself',
    fillcolor = '#FFDF70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3, 3, 0),
    theta = c(0, 262.5, 277.5, 0),
    fill = 'toself',
    fillcolor = '#B6FFB4',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,5)
      )
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    mode = 'lines'
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 1.5, 1.5, 0, 2.5, 2.5, 0),
    theta = c(0, 10, 25, 0, 205, 215, 0),
    fill = 'toself',
    fillcolor = '#709Bff',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3.5, 3.5, 0),
    theta = c(0, 55, 75, 0),
    fill = 'toself',
    fillcolor = '#E4FF87',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4.5, 4.5, 0, 4.5, 4.5, 0),
    theta = c(0, 100, 120, 0, 305, 320, 0),
    fill = 'toself',
    fillcolor = '#FFAA70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 4, 4, 0),
    theta = c(0, 165, 195, 0),
    fill = 'toself',
    fillcolor = '#FFDF70',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  add_trace(
    r = c(0, 3, 3, 0),
    theta = c(0, 262.5, 277.5, 0),
    fill = 'toself',
    fillcolor = '#B6FFB4',
    line = list(
      color = 'black'
    )
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,5)
      )
    ),
    showlegend = F
  )

fig
```

## Radar Charts
### Basic Radar Charts
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    fill = 'toself'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    ),
    showlegend = F
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    fill = 'toself'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    ),
    showlegend = F
  )

fig
```
### Multiple Trace Radar Charts
```r
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    fill = 'toself'
  ) 
fig <- fig %>%
  add_trace(
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group A'
  ) 
fig <- fig %>%
  add_trace(
    r = c(1.5, 10, 39, 31, 15, 1.5),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group B'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    )
  )

fig
```
```{r}
library(plotly)

fig <- plot_ly(
    type = 'scatterpolar',
    fill = 'toself'
  ) 
fig <- fig %>%
  add_trace(
    r = c(39, 28, 8, 7, 28, 39),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group A'
  ) 
fig <- fig %>%
  add_trace(
    r = c(1.5, 10, 39, 31, 15, 1.5),
    theta = c('A','B','C', 'D', 'E', 'A'),
    name = 'Group B'
  ) 
fig <- fig %>%
  layout(
    polar = list(
      radialaxis = list(
        visible = T,
        range = c(0,50)
      )
    )
  )

fig
```

